# Curate a master UKB event table {#curate-master-ukb-events-table}

This chapter gathers different types of fields available in the UKB assessment center data and generates a master event table. We also convert UKB defined special dates to normal dates. We take the following fields (details of which can be searched [here](https://biobank.ndph.ox.ac.uk/ukb/search.cgi)) from the UKB assessment center data:

- first occurrence fields in `first_occur_UKB.RDS`

- algorithmically defined outcome fields in `demog_UKB.RDS`:

  - `f.42000.0.0`
  - `f.42008.0.0`
  - `f.42010.0.0`
  - `f.42012.0.0`
  - `f.42006.0.0`
  - `f.42026.0.0`
  
- ICD10 code fields and their date fields in `ICD_UKB.RDS`:

  - starts with `f.41270` (ICD10 code)
  - starts with `f.41280` (ICD10 code event date)
  - starts with `f.40001` (ICD10 code primary death)
  - starts with `f.40002` (ICD10 code secondary death)
  - starts with `f.40000` (ICD10 code death date)
  
- OPCS4 code fields in `OPCS_procedures_UKB.RDS`:

  - starts with `f.41272` (OPCS4 code)
  - starts with `f.41282` (OPCS4 code event date)

- self-reported condition field in `demog_UKB.RDS`:

  - starts with `f.20002` (self-reported condition code)
  - starts with `f.20008` (self-reported condition code event date)
  
- self-reported operation field in `demog_UKB.RDS`:

  - starts with `f.20004` (self-reported operation code)
  - starts with `f.20010` (self-reported operation code event date)
  
In addition to these pre-defined fields, we define a custom field called `dr_self` which is a combination of the following fields:

  - `f.5901.0.0`
  - `f.5901.1.0`
  - `f.5901.2.0`
  - `f.5901.3.0`

These fields record age at which diabetic eye disease was diagnosed at four different time points. Using this information and the date of birth of a subject, we define the first occurrence event date for this customized outcome `dr_self`. Custom fields can be defined by users by combining multiple pre-defined fields. However, `dr_self` is the only custom outcome we define using the UKB assessment center data and included in the master event table.

When converting special dates to "normal" dates, we use the following mapping defined by UKB study:

- First occurrence, algorithimically defined outcome and OPC4 code event date fields

|Special date|Map|
|:------------|:---|
|1900-01-01|Missing|
|1901-01-01|Missing|
|2037-07-07|Missing||
|1902-02-02|DOB of a subject|
|1903-03-03|DOB of a subject|

- Self-reported condition and self-reported operation code event date fields:

|Special date|Map|
|:------------|:---|
|decimal date < 1900|Missing|

The conversion is carried out by the `cleandates()` function defined in `functions.R`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=F)
```

Load packages.
```{r, message = F}
library(tidyverse)
library(data.table)
library(lubridate)
source("functions.R")
```

Load reformatted raw UKB assessment data for generating a master UKB event table.
```{r}
firstoccurs <- readRDS("generated_data/first_occur_UKB.RDS")
ICD <- readRDS("generated_data/ICD_UKB.RDS")
procs <- readRDS("generated_data/OPCS_procedures_UKB.RDS")
demog <- readRDS("generated_data/demog_UKB.RDS")
```

Define the date of birth and gender
```{r}
demog <- demog %>% 
  rename(YOB = f.34.0.0) %>%
  rename(MOB = f.52.0.0) %>%
  mutate(DOB = lubridate::make_date(YOB, MOB)) %>%
  mutate(SEX = as.character(f.31.0.0))
```

```{r, echo=F, results='hide'}
#reduce datasets for memory management
#chunk not displayed
algo_outcome_fields <- paste0("f.",c("42000", "42008", "42010","42012","42006", "42026"),".0.0")
demog <- demog %>% select(f.eid, YOB, MOB, DOB, SEX, all_of(algo_outcome_fields), 
                          starts_with(c("f.5901.", "f.20002.", "f.20004.", "f.20008.", "f.20010.")))
ICD <- ICD %>% select(f.eid, 
                      starts_with(c("f.40000", "f.40001", "f.40002", "f.41271", "f.41281", "f.41280", "f.41270")))
procs <- procs %>% select(f.eid, starts_with(c("f.41272", "f.41282")))
gc()
```

## First occurrence event table and algorithmically defined outcome table

Convert first occurrence data into a long format.
```{r}
firstoccurs_long <- firstoccurs %>% 
  pivot_longer(-f.eid, names_to = "field", values_to = "event_dt", values_drop_na = T)
```

```{r, echo=F, results='hide'}
#Mem
rm(firstoccurs)
gc()
```

Obtain algorithmically defined outcomes from demographic dataset.
```{r}
algo_outcome_fields <- paste0("f.",c("42000", "42008", "42010","42012","42006", "42026"),".0.0")

algo_outcome_table_wide <- demog %>% select(f.eid,all_of(algo_outcome_fields))
algo_outcome_table_long <- algo_outcome_table_wide %>%  
  pivot_longer(-f.eid,names_to = "field",values_to = "event_dt", values_drop_na = T)
```

Merge algorithmically defined outcome fields in the demographic data with first occurrence data to produce outcome fields table.
```{r}
outcome_fields_table_long <- bind_rows(firstoccurs_long,algo_outcome_table_long)
```

## Custom defined outcome for DR

Define customized outcome for diabetes-related eye disease using demographic dataset. We use `5901` family of fields which include:

- `f.5901.0.0`
- `f.5901.1.0`
- `f.5901.2.0`
- `f.5901.3.0`

These fields record age at which DR was diagnosed at four different time points. Here are the steps for defining the first occurrence event data for this outcome:

1. Negative values that indicate incidence but unknown date of the event is converted to a numeric value `999`.
2. `dr_self = 1` if the event happened, or `0` otherwise.
3. `age_dr_self`: we take the youngest age at which a person was identified as having DR, and individuals with unknown date of this event is coded as `NA`.
4. If age is greater than `998`, then we are uncertain of when the outcome was actually identified, so they are converted back to `NA`.
5. `dr_self` is set to `NA` for individuals who were identified to have an outcome but with missing age when the outcome was identified.
6. To obtain the date of a DR event, we add DOB (decimal date) and age at which the outcome was identified.

```{r}
custom_outcome_fields_table_wide <- demog %>%
  mutate(f.5901.0.0 = replace(f.5901.0.0, which(f.5901.0.0 < 0), 999)) %>% 
  mutate(f.5901.1.0 = replace(f.5901.1.0, which(f.5901.1.0 < 0), 999)) %>% 
  mutate(f.5901.2.0 = replace(f.5901.2.0, which(f.5901.2.0 < 0), 999)) %>% 
  mutate(f.5901.3.0 = replace(f.5901.3.0, which(f.5901.3.0 < 0), 999)) %>% 
  mutate(dr_self = as.numeric(!is.na(f.5901.0.0) | !is.na(f.5901.1.0) | !is.na(f.5901.2.0) | !is.na(f.5901.3.0))) %>%
  mutate(age_dr_self = pmin(f.5901.0.0, f.5901.1.0, f.5901.2.0, f.5901.3.0, na.rm=T)) %>%
  mutate(age_dr_self = replace(age_dr_self, which(age_dr_self > 998), NA)) %>%
  mutate(dr_self = replace(dr_self, which(is.na(age_dr_self) & dr_self==1), NA)) %>%
  mutate(date_dr_self = as.Date(date_decimal(decimal_date(DOB) + age_dr_self))) %>% 
  select(f.eid,date_dr_self)

custom_outcome_fields_table_long <- 
  custom_outcome_fields_table_wide %>% 
  pivot_longer(-f.eid, names_to = "field", values_to = "event_dt", values_drop_na = T)

custom_outcome_fields_table_long <-
  custom_outcome_fields_table_long %>% 
  mutate(field = ifelse(field == "date_dr_self", "dr_self", field))
```

## ICD 9 & 10

Obtain relevant fields.
```{r}
ICD9_codes <- ICD %>% select(f.eid,starts_with('f.41271')) %>% arrange(f.eid) %>% data.frame()
ICD9_dates <- ICD %>% select(f.eid,starts_with('f.41281.')) %>% arrange(f.eid) %>% data.frame()

ICD10_codes <- ICD %>% select(f.eid,starts_with('f.41270')) %>% arrange(f.eid) %>% data.frame()
ICD10_dates <- ICD %>% select(f.eid,starts_with('f.41280.')) %>% arrange(f.eid) %>% data.frame() 

ICD10_codes_primary_death <- ICD %>% select(f.eid,starts_with('f.40001')) %>% arrange(f.eid) %>% data.frame()
ICD10_codes_secondary_death <- ICD %>% select(f.eid,starts_with('f.40002')) %>% arrange(f.eid) %>% data.frame()
ICD10_death_date <- ICD %>% select(f.eid, f.40000.0.0) %>% arrange(f.eid) %>% data.frame()
```

```{r, echo=F, results='hide'}
#mem
rm(ICD)
gc()
```

Merge ICD10 primary death and death date tables into a long format.
```{r}
ICD10_codes_primary_death_long <-
  left_join(ICD10_codes_primary_death %>% pivot_longer(cols=-f.eid),ICD10_death_date,by="f.eid") %>% 
  select(-name) %>% rename(code=value,event_dt=`f.40000.0.0`) %>% mutate(type="ICD10_death_primary") %>% distinct()
```

Merge ICD10 secondary death and death date tables into a long format.
```{r}
ICD10_codes_secondary_death_long <-
  left_join(ICD10_codes_secondary_death %>% pivot_longer(cols=-f.eid),ICD10_death_date,by="f.eid") %>% 
  select(-name) %>% rename(code=value,event_dt=`f.40000.0.0`) %>% mutate(type="ICD10_death_secondary") %>% distinct()
```

Merge ICD10 codes and their event dates tables into a long format.
```{r, warning=F, message=F}
ICD10_codes_long <- merge_long("f.41270","f.41280",ICD10_codes,ICD10_dates,"ICD10") %>% distinct()
```

Merge ICD9 codes and their event dates tables into a long format.
```{r}
ICD9_codes <- ICD9_codes %>% mutate_at(vars(f.41271.0.0:f.41271.0.46), as.character)
ICD9_codes_long <- merge_long("f.41271","f.41281",ICD9_codes,ICD9_dates,"ICD9") %>% distinct()
```

Merge all of the ICD event tables with dates, and then remove observations with no codes (`code == NA`). 
```{r}
ICD_codes_full <- do.call(rbind,list(ICD10_codes_long,ICD9_codes_long,ICD10_codes_primary_death_long,ICD10_codes_secondary_death_long)) %>% filter(!is.na(code))
```

## OPCS4

OPCS4 event table is procedure codes from hospital admission data. Obtain relevant fields.
```{r}
OPCS4 <- procs %>% 
  select(c(f.eid, starts_with('f.41272.'))) %>% 
  mutate_if(is.factor, as.character)  %>% 
  arrange(f.eid) %>% 
  data.frame()

OPCS4dates <- procs %>% 
  select(c(f.eid, starts_with('f.41282.'))) %>% 
  arrange(f.eid) %>% 
  data.frame()

OPCS4_codes_long <- 
  merge_long("f.41272","f.41282",OPCS4,OPCS4dates,"OPCS4") %>% distinct() %>% filter(!is.na(code))
```

## Self-reported conditions

Obatin relevant fields.
```{r, warning=F}
selfrep_codes <- demog %>%
  select(c(f.eid, starts_with("f.20002."))) %>%
  arrange(f.eid) %>% 
  data.frame() # can't be data.table

selfrep_dates <- demog %>%
  select(c(f.eid, starts_with("f.20008."))) %>%
  data.frame() %>% 
  mutate_all(funs(replace(., .<1900, NA))) %>%
  mutate_at(vars(starts_with("f.20008.")), .funs = list(~ lubridate::date_decimal(.))) %>%
  mutate_at(vars(starts_with("f.20008.")), .funs = list(~ as.Date(.))) %>%
  arrange(f.eid)

selfrep_codes_long <- merge_long("f.20002","f.20008",selfrep_codes,selfrep_dates,"selfrep") %>% 
  distinct() %>% filter(!is.na(code))
```


## Self-reported operations

Obatin relevant fields.
```{r}
selfrep_op_codes <- demog %>%
  select(c(f.eid, starts_with("f.20004."))) %>%
  arrange(f.eid) %>% 
  data.frame()

selfrep_op_dates <- demog %>%
  select(c(f.eid, starts_with("f.20010."))) %>%
  data.frame() %>%
  mutate_all(funs(replace(., .<1900, NA))) %>%
  mutate_at(vars(starts_with("f.20010.")), .funs = list(~ lubridate::date_decimal(.))) %>%
  mutate_at(vars(starts_with("f.20010.")), .funs = list(~ as.Date(.))) %>%
  arrange(f.eid) 

selfrep_op_codes_long <- merge_long("f.20004","f.20010",selfrep_op_codes,selfrep_op_dates,"selfrep_op") %>% 
  distinct() %>% filter(!is.na(code))
```

```{r, echo=F, results='hide'}
#mem
rm(procs)
gc()
```

## Standardize and merge event tables

The following event tables are merged to produce a master event table:

- Outcome field event table (combination of first occurrence event data and algorithmically defined event data from demographic table)
- ICD10 code event table
- OPCS4 code event table
- Self-reported condition code event table
- Self-reported operation code event table
- Custom defined event table for DR

```{r}
# Standardize
outcome_fields_table_long <- 
  outcome_fields_table_long %>% mutate(type = "outcome_fields") %>% rename(key=field)

ICD_codes_full <- 
  ICD_codes_full %>% mutate(code=as.character(code)) %>% rename(key=code)

OPCS4_codes_long <- 
  OPCS4_codes_long %>% rename(key=code)

selfrep_codes_long <- 
  selfrep_codes_long %>% mutate(code=as.character(code)) %>% rename(key=code)

selfrep_op_codes_long <- 
  selfrep_op_codes_long %>% mutate(code=as.character(code)) %>% rename(key=code)

custom_outcome_fields_table_long <-
  custom_outcome_fields_table_long %>% mutate(type = "custom_fields") %>% rename(key = field)

event_tab <- bind_rows(list(outcome_fields_table_long,ICD_codes_full,
                            OPCS4_codes_long,selfrep_codes_long,selfrep_op_codes_long,
                            custom_outcome_fields_table_long))
```

Filter out events where the date of event is missing.
```{r}
event_tab <- event_tab %>% filter(!is.na(event_dt))
```

## Convert UKB special dates into "normal" dates

Convert special dates to normal dates and then filter out any event with an unknown event date. Note that special dates in self-reported and self-reported operation and custom defined outcome table have already been converted in the above. We convert special dates present in the following event table types: `outcome_fields`, `ICD10`, `ICD10_death_primary`, `ICD10_death_secondary` and `OPCS4`. 
```{r}
demog_dob <- demog %>% select(f.eid, DOB)
event_tab <- event_tab %>% left_join(demog_dob, by = "f.eid")
event_tab$event_dt <- cleandates(event_tab$event_dt,event_tab$DOB)
event_tab <- event_tab %>% select(-DOB) %>% filter(!is.na(event_dt))
```

Save all UKB events table. Note that this table includes subjects whose genetic and reported sex do not match.
```{r}
saveRDS(event_tab,"generated_data/pre_all_ukb_events_tab.RDS")
```




