# Curate all UKB events table {#curate-all-ukb_events-table}

UK Biobank assessment center data (or in short UKB data) are in the wide format where each field represents some event. We identify what events these fields represent in [here](https://biobank.ndph.ox.ac.uk/ukb/search.cgi).

* Generated file: `pre_all_ukb_events_tab.RDS`

We use the following types of fields

- first occurrence fields in the first occurrence event table

- algorithmically defined outcome fields:
  - f.42000.0.0
  - f.42008.0.0
  - f.42010.0.0
  - f.42012.0.0
  - f.42006.0.0
  - f.42026.0.0
  
- ICD10 code fields and their date fields:
  - starts with f.41270 (ICD10 code)
  - starts with f.41280 (ICD10 code event date)
  - starts with f.40001 (ICD10 code primary death)
  - starts with f.40002 (ICD10 code secondary death)
  - starts with f.40000 (ICD10 code death date)
  
- OPCS4 code fields
  - starts with f.41272 (OPCS4 code)
  - starts with f.41282 (OPCS4 code event date)

- Self-reported condition field
  - starts with f.20002 (self-reported condition code)
  - starts with f.20008 (self-reported condition code event date)
  
- Self-reported operation field
  - starts with f.20004 (self-reported operation code)
  - starts with f.20010 (self-reported operation code event date)


Once we have collected the tables containing these fields, we convert them into a long format and merge them together to create a master UKB event table. We can then simply search different fields and/or code patterns in the master event table to curate an event table of interest. The fields and/or codes we can specify to search in the master event table are:

- first occurrence field patterns
- algorithmically defined outcome field patterns
- ICD10 code patterns
- OPCS4 code patterns
- Self-reported condition code patterns
- Self-reported operation code patterns
- Custom field patterns

Note that custom field is user defined field that combines multiple fields. For example, `dr_self` field considers the following fields:

- f.5901.0.0
- f.5901.1.0
- f.5901.2.0
- f.5901.3.0

These fields record age at which diabetic eye disease was diagnosed at four different time points. Using this information and the date of birth of subjects we define the first occurrence event date for this outcome. Currently, `dr_self` is the only custom field that exists in the master event table.

We also converted the special dates using predefined mapping. The special dates and their mappings:

- For first occurrence, algorithimically defined outcome and OPC4 code event date fields:

|Special date|Map|
|------------|---|
|1900-01-01|Missing|
|1901-01-01|Missing|
|2037-07-07|Missing||
|1902-02-02|DOB of a subject|
|1903-03-03|DOB of a subject|

- For self-reported condition and self-reported operation code event date fields:

|Special date|Map|
|------------|---|
|decimal date < 1900|Missing|

Generated file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=T)
```

```{r, message = F}
library(tidyverse)
library(data.table)
library(lubridate)
source("functions.R")
```

This step compiles all UKB events table that will be used in phenotyping diabetes and diabetes-related complications. 

Here, we accomplish:

- Gather different types of event tables and generate a master event table in a long format
- Convert UKB defined special dates to normal dates for outcome event dates

Load reformatted raw UKB assessment data needed generate a master event table
```{r}
firstoccurs <- readRDS("generated_data/first_occur_UKB.RDS")
ICD <- readRDS("generated_data/ICD_UKB.RDS")
procs <- readRDS("generated_data/OPCS_procedures_UKB.RDS")
demog <- readRDS("generated_data/demog_UKB.RDS")
```

Define the date of birth and gender
```{r}
demog <- demog %>% 
  rename(YOB = f.34.0.0) %>%
  rename(MOB = f.52.0.0) %>%
  mutate(DOB = lubridate::make_date(YOB, MOB)) %>%
  mutate(SEX = as.character(f.31.0.0))
```

```{r, echo=F, results='hide'}
#reduce datasets for memory management
#chunk not displayed
algo_outcome_fields <- paste0("f.",c("42000", "42008", "42010","42012","42006", "42026"),".0.0")
demog <- demog %>% select(f.eid, YOB, MOB, DOB, SEX, all_of(algo_outcome_fields), 
                          starts_with(c("f.5901.", "f.20002.", "f.20004.", "f.20008.", "f.20010.")))
ICD <- ICD %>% select(f.eid, 
                      starts_with(c("f.40000", "f.40001", "f.40002", "f.41271", "f.41281", "f.41280", "f.41270")))
procs <- procs %>% select(f.eid, starts_with(c("f.41272", "f.41282")))
gc()
```


## First occurrence event table and algorithmically defined outcome table

Convert first occurrence data into a long format
```{r}
firstoccurs_long <- firstoccurs %>% 
  pivot_longer(-f.eid, names_to = "field", values_to = "event_dt", values_drop_na = T)
```

```{r, echo=F, results='hide'}
#Mem
rm(firstoccurs)
gc()
```

Obtain algorithmically defined outcomes from demographic dataset.
```{r}
algo_outcome_fields <- paste0("f.",c("42000", "42008", "42010","42012","42006", "42026"),".0.0")

algo_outcome_table_wide <- demog %>% select(f.eid,all_of(algo_outcome_fields))
algo_outcome_table_long <- algo_outcome_table_wide %>%  
  pivot_longer(-f.eid,names_to = "field",values_to = "event_dt", values_drop_na = T)
```

Merge algorithmically defined outcome fields in the demographic data with first occurrence data to produce outcome fields table
```{r}
outcome_fields_table_long <- bind_rows(firstoccurs_long,algo_outcome_table_long)
```

## Custom defined outcome for DR

Define customized outcome of diabetes-related eye disease using demographic dataset. We use fields 5901 family which includes:

- f.5901.0.0
- f.5901.1.0
- f.5901.2.0
- f.5901.3.0

These fields record age at which DR was diagnosed at four different time points. Here are the steps for defining the first occurrence event data for this outcome :

1. Negative values that indicate incidence but unknown date of the event is converted to a numeric value 999.
2. `dr_self` = 1 if the event happened, or 0 otherwise.
3. `age_dr_self`: We take the youngest age at which a person was identified as having DR, and individuals with unknown date of this event is coded as NA.
4. If age is greater than 998, then we are uncertain of when the outcome was actually identified, so they are converted back to NA.
5. `dr_self` is set to NA for individuals who were identified to have an outcome but with missing age when the outcome was identified.
6. We add DOB and age when the outcome was identified to obtain the date of an event.

```{r}
custom_outcome_fields_table_wide <- demog %>%
  mutate(f.5901.0.0 = replace(f.5901.0.0, which(f.5901.0.0 < 0), 999)) %>% 
  mutate(f.5901.1.0 = replace(f.5901.1.0, which(f.5901.1.0 < 0), 999)) %>% 
  mutate(f.5901.2.0 = replace(f.5901.2.0, which(f.5901.2.0 < 0), 999)) %>% 
  mutate(f.5901.3.0 = replace(f.5901.3.0, which(f.5901.3.0 < 0), 999)) %>% 
  mutate(dr_self = as.numeric(!is.na(f.5901.0.0) | !is.na(f.5901.1.0) | !is.na(f.5901.2.0) | !is.na(f.5901.3.0))) %>%
  mutate(age_dr_self = pmin(f.5901.0.0, f.5901.1.0, f.5901.2.0, f.5901.3.0, na.rm=T)) %>%
  mutate(age_dr_self = replace(age_dr_self, which(age_dr_self > 998), NA)) %>%
  mutate(dr_self = replace(dr_self, which(is.na(age_dr_self) & dr_self==1), NA)) %>%
  mutate(date_dr_self = as.Date(date_decimal(decimal_date(DOB) + age_dr_self))) %>% 
  select(f.eid,date_dr_self)

custom_outcome_fields_table_long <- 
  custom_outcome_fields_table_wide %>% 
  pivot_longer(-f.eid, names_to = "field", values_to = "event_dt", values_drop_na = T)

custom_outcome_fields_table_long <-
  custom_outcome_fields_table_long %>% 
  mutate(field = ifelse(field == "date_dr_self", "dr_self", field))
```

## ICD 9 & 10

Obtain relevant fields
```{r}
ICD9_codes <- ICD %>% select(f.eid,starts_with('f.41271')) %>% arrange(f.eid) %>% data.frame()
ICD9_dates <- ICD %>% select(f.eid,starts_with('f.41281.')) %>% arrange(f.eid) %>% data.frame()

ICD10_codes <- ICD %>% select(f.eid,starts_with('f.41270')) %>% arrange(f.eid) %>% data.frame()
ICD10_dates <- ICD %>% select(f.eid,starts_with('f.41280.')) %>% arrange(f.eid) %>% data.frame() 

ICD10_codes_primary_death <- ICD %>% select(f.eid,starts_with('f.40001')) %>% arrange(f.eid) %>% data.frame()
ICD10_codes_secondary_death <- ICD %>% select(f.eid,starts_with('f.40002')) %>% arrange(f.eid) %>% data.frame()
ICD10_death_date <- ICD %>% select(f.eid, f.40000.0.0) %>% arrange(f.eid) %>% data.frame()
```

```{r, echo=F, results='hide'}
#mem
rm(ICD)
gc()
```

Merge ICD10 primary death and death date tables into a long format
```{r}
ICD10_codes_primary_death_long <-
  left_join(ICD10_codes_primary_death %>% pivot_longer(cols=-f.eid),ICD10_death_date,by="f.eid") %>% 
  select(-name) %>% rename(code=value,event_dt=`f.40000.0.0`) %>% mutate(type="ICD10_death_primary") %>% distinct()
```

Merge ICD10 secondary death and death date tables into a long format
```{r}
ICD10_codes_secondary_death_long <-
  left_join(ICD10_codes_secondary_death %>% pivot_longer(cols=-f.eid),ICD10_death_date,by="f.eid") %>% 
  select(-name) %>% rename(code=value,event_dt=`f.40000.0.0`) %>% mutate(type="ICD10_death_secondary") %>% distinct()
```

Merge ICD10 codes and their event dates tables into a long format
```{r, warning=F, message=F}
ICD10_codes_long <- merge_long("f.41270","f.41280",ICD10_codes,ICD10_dates,"ICD10") %>% distinct()
```

Merge ICD9 codes and their event dates tables into a long format
```{r}
ICD9_codes <- ICD9_codes %>% mutate_at(vars(f.41271.0.0:f.41271.0.46), as.character)
ICD9_codes_long <- merge_long("f.41271","f.41281",ICD9_codes,ICD9_dates,"ICD9") %>% distinct()
```


Merge all of the ICD event tables with dates, and then remove observations with no codes (code == NA). 
```{r}
ICD_codes_full <- do.call(rbind,list(ICD10_codes_long,ICD9_codes_long,ICD10_codes_primary_death_long,ICD10_codes_secondary_death_long)) %>% filter(!is.na(code))
```

## OPCS4

OPCS4 event table is procedure codes from hospital admission data.
```{r}
OPCS4 <- procs %>% 
  select(c(f.eid, starts_with('f.41272.'))) %>% 
  mutate_if(is.factor, as.character)  %>% 
  arrange(f.eid) %>% 
  data.frame()

OPCS4dates <- procs %>% 
  select(c(f.eid, starts_with('f.41282.'))) %>% 
  arrange(f.eid) %>% 
  data.frame()

OPCS4_codes_long <- 
  merge_long("f.41272","f.41282",OPCS4,OPCS4dates,"OPCS4") %>% distinct() %>% filter(!is.na(code))
```

## Medical conditions self-report UKB

```{r, warning=F}
selfrep_codes <- demog %>%
  select(c(f.eid, starts_with("f.20002."))) %>%
  arrange(f.eid) %>% 
  data.frame() # can't be data.table

selfrep_dates <- demog %>%
  select(c(f.eid, starts_with("f.20008."))) %>%
  data.frame() %>% 
  mutate_all(funs(replace(., .<1900, NA))) %>% # it looks like this step got rid of unknown date of an outcome event
  mutate_at(vars(starts_with("f.20008.")), .funs = list(~ lubridate::date_decimal(.))) %>%
  mutate_at(vars(starts_with("f.20008.")), .funs = list(~ as.Date(.))) %>%
  arrange(f.eid)

selfrep_codes_long <- merge_long("f.20002","f.20008",selfrep_codes,selfrep_dates,"selfrep") %>% 
  distinct() %>% filter(!is.na(code))
```


## Self-reported operations UKB

```{r}
selfrep_op_codes <- demog %>%
  select(c(f.eid, starts_with("f.20004."))) %>%
  arrange(f.eid) %>% 
  data.frame()

selfrep_op_dates <- demog %>%
  select(c(f.eid, starts_with("f.20010."))) %>%
  data.frame() %>%
  mutate_all(funs(replace(., .<1900, NA))) %>% # it looks like this step got rid of unknown date of an outcome event
  mutate_at(vars(starts_with("f.20010.")), .funs = list(~ lubridate::date_decimal(.))) %>%
  mutate_at(vars(starts_with("f.20010.")), .funs = list(~ as.Date(.))) %>%
  arrange(f.eid) 

selfrep_op_codes_long <- merge_long("f.20004","f.20010",selfrep_op_codes,selfrep_op_dates,"selfrep_op") %>% 
  distinct() %>% filter(!is.na(code))
```

```{r, echo=F, results='hide'}
#mem
rm(procs)
gc()
```

## Standardize and merge event tables

The following event tables are merged to produce a master event table of UKB assessment data:

- Outcome field event table (combination of first occurrence event data and algorithmically defined event data from demographic table)
- ICD10 code event table
- OPCS4 code event table
- Self-reported condition code event table
- Self-reported operation code event table
- Custom defined event table for DR

```{r}
# Standardize
outcome_fields_table_long <- 
  outcome_fields_table_long %>% mutate(type = "outcome_fields") %>% rename(key=field)

ICD_codes_full <- 
  ICD_codes_full %>% mutate(code=as.character(code)) %>% rename(key=code)

OPCS4_codes_long <- 
  OPCS4_codes_long %>% rename(key=code)

selfrep_codes_long <- 
  selfrep_codes_long %>% mutate(code=as.character(code)) %>% rename(key=code)

selfrep_op_codes_long <- 
  selfrep_op_codes_long %>% mutate(code=as.character(code)) %>% rename(key=code)

custom_outcome_fields_table_long <-
  custom_outcome_fields_table_long %>% mutate(type = "custom_fields") %>% rename(key = field)

event_tab <- bind_rows(list(outcome_fields_table_long,ICD_codes_full,
                            OPCS4_codes_long,selfrep_codes_long,selfrep_op_codes_long,
                            custom_outcome_fields_table_long))
```

Filter out events where the date of event is missing
```{r}
event_tab <- event_tab %>% filter(!is.na(event_dt))
```

## Convert UKB special dates into "normal" dates

Convert special dates to normal dates, and then filter out any event with an unknown event date. Note that special dates in self-reported and self-reported operation and custom defined outcome table have already been converted in the above. We convert special dates present in the following event table types: `outcome_fields`, `ICD10`, `ICD10_death_primary`, `ICD10_death_secondary` and `OPCS4`. 
```{r}
demog_dob <- demog %>% select(f.eid, DOB)
event_tab <- event_tab %>% left_join(demog_dob, by = "f.eid")
event_tab$event_dt <- cleandates(event_tab$event_dt,event_tab$DOB)
event_tab <- event_tab %>% select(-DOB) %>% filter(!is.na(event_dt))
```

Save all UKB events table. Note that this table includes subjects with genetic and reported sex mismatch
```{r}
saveRDS(event_tab,"generated_data/pre_all_ukb_events_tab.RDS")
```




