# Create a master code dictionary for primary care data {#make-code-dictionary}

To phenotype a clinical event using primary care data, we need to identify a set of codes that represents the outcome of interest. To identify these codes, we first curate a master code dictionary which contains all of the codes and descriptions associated with primary care data. The master dictionary is a combination of the following types of code dictionaries, which we pull together to generate a master dictionary.:

- read v2
- read v3
- TPP

The purpose of curating the master dictionary is so that we can subset the master dictionary and create a outcome-specific dictionary. For example, for phenotyping diabetes subjects, we would identify a certain set of codes that are known to indicate diabetes, and also search for a certain set of keywords in the descriptions of the codes since each code has multiple descriptions tied to it. Then, using diabetes-specific dictionary, we could identify subjects with diabetes from the PCP data. Curating outcome-specific dictionary is carried out in  \@ref(make-outcome-specific-code-dictionary).

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, eval = F)
```

```{r}
library(data.table)
library(tidyverse)
```

## Read v2

Import Read v2 code dictionary data:
```{r}
read2_dict_drug <- fread("raw_data/read_v2_drugs.csv")
read2_dict_nondrug <- fread("raw_data/read_v2_V2.csv")
```

Combine drug and nondrug Read v2 dictionaries. Keep only read_code and term_description fields.
```{r}
read2_dict <- full_join(read2_dict_drug,read2_dict_nondrug) %>%
  select(read_code,term_description)
```

## Read v3

Import Read v3 code dictionary data. Read v3 code dictionary includes code dictionary for drugs. "read_v3_V2.csv" is missing some fields. So, import "read_v3.csv" instead.
```{r}
read3_dict <- fread("raw_data/read_v3_V2.csv")
```

Note that Read v3 dictionary has identical codes that start with period and without, for example, .9m05 vs 9m05. In the UKB database, the one starting with '.' is never used. So, filter out codes that start with a lower case letter because these are the drug codes for Read v3 (CTV3) that match Read v2.
```{r}
read3_dict <- read3_dict %>% 
  select(read_code, term_description) %>%
  filter(!(grepl('^\\.', read_code)))
```

Some Read v3 descriptions merely points to another Read v3 code rather than actual description, for example, the term description says "See XE0md." So, we will replace this kind of term description with the actual description of the code. First, make look up table.
```{r}
lookup <- read3_dict$term_description
names(lookup) <- read3_dict$read_code
```

Check whether all the codes in "See [code]" description appear in the list of Read v3 codes. There are a couple of codes with description "See [Code]" that are not found keys of the lookup table. The descriptions of these codes will be replaced with NA.
```{r}
temp <- read3_dict %>%
  mutate(code_fillin = ifelse(grepl("^See ", term_description), gsub("See ", "", term_description), NA)) 
temp2 <- temp %>% filter(!is.na(code_fillin)) %>% .$code_fillin
temp2[!(temp2 %in% names(lookup))]
```

Replace term description of the form "See [code]" with the actual description of the code.
```{r}
read3_dict <- read3_dict %>%
  mutate(code_fillin = ifelse(grepl("^See ", term_description), gsub("See ", "", term_description), NA)) %>%
  mutate(term_description = ifelse(is.na(code_fillin), term_description, lookup[code_fillin])) %>%
  select(-code_fillin)
```

We check whether Read v2 and Read v3 codes overlap. It seems all codes in Read v2 dictionary appear in Read v3 dictionary except for the following:
```{r}
read2_dict$read_code[!(read2_dict$read_code %in% read3_dict$read_code)]
```

## TPP local codes

Import TPP local codes which are also used in primary care data. These begin with 'Y' (Note: `fread` automatically resolved the parsing error that `read_delim` could not do, so we used `fread` to import the TPP local code data).
```{r}
tpp_local <- fread("raw_data/tpp_local.txt")
```

TPP codes do not appear in Read v2
```{r}
tpp_local %>% filter(code %in% read2_dict$read_code)
```

A couple of TPP codes appears in Read v3:
```{r}
tpp_local %>% filter(code %in% read3_dict$read_code)
```

## Merge Read v2, Read v3 and TPP dictionaries

Note that this merged dictionary is intended to cover all the descriptions used in Read v2, Read v3 and TPP so that we do not miss a description when are phenotyping an outcome later. So, the merged dictionary have one-to-many mapping from code to descriptions which is intentional and what we want.
```{r}
full_dict <- 
  full_join(read2_dict %>% rename(code = read_code) %>% mutate(terminology="read2"),
            read3_dict %>% rename(code = read_code) %>% mutate(terminology="read3")) %>%
  full_join(tpp_local %>% mutate(terminology="read3", terminology_note = "TPP Local Code")) %>%
  distinct() 
```

Save the full code dictionary for primary care events
```{r}
saveRDS(full_dict,"generated_data/full_dict.RDS")
```


