# Make outcome-specific code dictionary {#make-outcome-specific-code-dictionary}

In order to obtain outcome-specific dictionary, we want to first specify the following:

- inclusion codes or code patterns
- exclusion codes or code patterns
- inclusion keywords or keyword patterns
- exclusion keywords or keyword patterns

Using these, we subset the master dictionary (`full_dict.RDS`) to generate outcome-specific dictionary. The following are the event tables that require the use of PCP data and thus require outcome-specific dictionary:

- Diabetes event table
- Diabetic eye disease table for capturing diabetic eye disease cases when curating time-to-event data
- Non-diabetic eye disease table for excluding controls for time-to-event data curation
- Kidney disease case event table for capturing diabetic kidney disease cases when curating time-to-event data
- Kidney disease event table for excluding controls for time-to-event data curation

We achieve the following in this step:

- Generate diabetes specific dictionary 
- Generate diabetic eye disease dictionary
- Generate non-diabetic eye disease dictionary
- Generate kidney disease case dictionary
- Generate kidney disease control exclusion event dictionary

Input and output files:

- Input (from `raw_data` folder): 

  - `dm_code_lists` folder:
    - `opensafely-diabetes-2020-04-15.csv`
    - `opensafely-diabetes-exeter-group-2020-07-06.csv`
    - `opensafely-type-1-diabetes-2020-06-29.csv`
    - `opensafely-type-2-diabetes-2020-06-29.csv`
    - `read_diabetes_caliber.txt`
    - `read_diabetes_expanded_caliber.txt`
    - `read_diabetescomplications_caliber.txt`
  
  - `kd_code_lists` folder:
    - `opensafely-chronic-kidney-disease-2020-04-14.csv`
    - `opensafely-dialysis-2020-07-16.csv`
    - `opensafely-kidney-transplant-2020-07-15.csv`
  
- Output:

  - `dm_dict.RDS`
  - `dm_eye_disease_dict.RDS`
  - `nondm_eye_disease_dict.RDS`
  - `kidney_disease_case_dict.RDS`
  - `kidney_disease_control_exclusion_dict.RDS`
  
## Prepare diabetes code dictionary

### Define keywords

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, eval = F)
```

```{r}
library(tidyverse)
library(data.table)
```

Global exclusion keywords. Codes that relate to the absence of a condition, screening for a condition, normal result for a condition, family history of a condition, various administrative statuses, and more, are not wanted. I exclude them from all of the following searches.
```{r}
global_exclusion_keyword_patterns <- " normal| absent|inspection|examination|evaluation|referral|^no |^O/E- no |^O/E - no |excluded|not seen|seen by|not examined|did not attend|assessable|offered|^trauma| trauma|malnutrition|screening for|^FH|family history|interpret|under care|refused|unsuitable|at risk|resolved|disease screening|not indicated|letter|excepted|exception|fear of|education|sharing consent|tumour|pregnancy|gestational|diabetic child|complication of care|steriod-induced|lymphoma|cytomegalovirus|rubella|blastoma|histoplasm|toxoplasm|neoplasm|herpes|tubercul|pediculus| observation$|injury|streptococcal"
```

DM specific exclusion keywords
```{r}
dm_specific_exclusion_keyword_patterns <- "|serum|antibody|remission|relative|association member|high risk of diabetes|suspected diabetes|non-diabetes|breath test|questionnaire|risk score|category score|risk calculator|inhibitor function|C-peptide level|factor binding protein 3|stress test|Insulin tolerance test|Urine screening test|X-ray|NHS Diabetes Prevention|Provision of diabetes clinical summary|diabetes mellitus screen|leaflet given|declined|C1-esterase|pituitary|helicobacter|ineligible|invite|invitation|insulinoma|steroid|secondary diabetes|pre-diabet|prediabet|insipid|provision of written information|not required|national audit|diabetes screen|renal diabetes|non-diabet|^diabetic nurse$|^Diabetic liaison nurse$|jamaica|secondary pancreatic diabetes|driving|neonatal|Addison|PABA test|growth factor|Plasma insulin level|key contact|eligibiliby|CHA2DS2|Professional judgement|Diabetes mellitus: no|non diabetic|information prescription|mother has|preg.|bronzed|Diabetes dietitian|Urine Ketone Test|deleted|refuse|gastropathy|pneumon|Frequency of hypoglycaem|^insulin level$|drug-induced|drug induced"
```

DM exclusion keywords
```{r}
dm_exclusion_keyword_patterns <- paste0(global_exclusion_keyword_patterns,dm_specific_exclusion_keyword_patterns)
```

Inclusion keywords
```{r}
dm_inclusion_keyword_patterns <- "diabetic|diabetes|diabeto|insulin|hyperglyc|hypoglyc|glycemic control"
```

### Define codes

DM exclusion codes
```{r}
dm_exclusion_codes <- c("ZV653", "C3760", "J4z0", "Y3045", "7L1L2", "Y0015", "X789v", "Y7ITk", "Y2200", "42c..", "42W..", "42WZ.","66Ae.", "66Ae0","66AF.","C1…","XaCET","XaCEU", "XaCEV")
dm_exclusion_code_patterns <- paste(c("^42W","^42c"),collapse = '|')
```

Import all the DM code lists
```{r}
dm_codelist <- fread("raw_data/dm_code_lists/opensafely-diabetes-2020-04-15.csv") %>% 
  full_join(fread("raw_data/dm_code_lists/opensafely-type-1-diabetes-2020-06-29.csv") 
            %>% mutate(Category=1)) %>%
  full_join(fread("raw_data/dm_code_lists/opensafely-type-2-diabetes-2020-06-29.csv") 
            %>% mutate(Category=2)) %>%
  full_join(fread("raw_data/dm_code_lists/opensafely-diabetes-exeter-group-2020-07-06.csv") 
            %>% dplyr::rename(CTV3PreferredTermDesc = ctvterm)) %>%
  dplyr::rename(term_description = CTV3PreferredTermDesc) %>%
  dplyr::rename(code = CTV3ID) %>%
  mutate(Category = as.character(Category)) %>%
  full_join(
    fread("raw_data/dm_code_lists/read_diabetescomplications_caliber.txt") %>%
    full_join(fread("raw_data/dm_code_lists/read_diabetes_expanded_caliber.txt")) %>%
    full_join(fread("raw_data/dm_code_lists/read_diabetes_caliber.txt")) %>%
      dplyr::rename(code = Clinical_code) %>%
      dplyr::rename(term_description = Clinical_term) %>%
      dplyr::rename(Category = `Category_(code)`)
    )
```

Define inclusion codes
```{r}
dm_inclusion_codes <- dm_codelist$code
```

Define additional DM code patterns we want
```{r}
dm_inclusion_code_patterns <- paste(c("^66A","^C10","^F420"),collapse = '|')
```

### Create diabetes code dictionary

Import full dictionary
```{r}
full_dict <- readRDS("generated_data/full_dict.RDS")
```

Create diabetes code dictionary
```{r}
dm_dict <- full_dict %>% 
  filter(grepl(dm_inclusion_keyword_patterns, term_description, ignore.case = T)|
           code %in% dm_inclusion_codes|
           grepl(dm_inclusion_code_patterns, code)) %>%
  filter(!grepl(dm_exclusion_keyword_patterns, term_description, ignore.case = T),
         !(code %in% dm_exclusion_codes),
         !grepl(dm_exclusion_code_patterns, code))
```

Save diabetes code dictionary
```{r,include=T}
saveRDS(dm_dict,"generated_data/dm_dict.RDS")
```

## Prepare diabetic eye disease code dictionary

### Define keywords

Inclusion keywords
```{r}
inclusion_keyword_patterns <- 
  "macul|retin|ophthalmic manifestation|ophthalmic complication|cataract|glaucoma|eye disease|diabetic iritis"
```

Inclusion keywords for case
```{r}
inclusion_keyword_patterns_case <- "diabetic|diabetes|diabet retinopathy"
```

Global exclusion keywords
```{r}
global_exclusion_keyword_patterns <- 
  "normal| absent|inspection|examination|evaluation|referral|^no |^O/E- no |^O/E - no |excluded|not seen|seen by|not examined|did not attend|assessable|offered|^trauma| trauma|malnutrition|screening for|^FH|family history|interpret|under care|refused|unsuitable|at risk|resolved|disease screening|not indicated|letter|excepted|exception|fear of|education|sharing consent|tumour|pregnancy|gestational|diabetic child|complication of care|steriod-induced|lymphoma|cytomegalovirus|rubella|blastoma|histoplasm|toxoplasm|neoplasm|herpes|tubercul|pediculus| observation$|injury|streptococcal"
```

Exclusion keywords
```{r}
exclusion_keyword_patterns <-
  " no maculopathy|^normal |translocation|cataract screen|retinol|incretin|diabetic retinopathy screening|retinal screening|steroid|drug induced|ratiation induced|branch of|vocational asses|electroretinography|buckling|tamponade|biopsy|retinopexy|cretin|glaucoma screen|pigment|fundoscopy|migraine|Commotio retinae|angiography|melanocytic macule|melanotic macule|acute retinal necrosis|radiation retinopathy|solar retinopathy|retinal dialysis|toxic maculopathy|eruption|retinacul|Acitretin|tretinoin|alopecia|branch of retinal artery$|rash"

exclusion_keywords <- c("O/E - retina", "Retinal photography", "Retinoscopy", 
                                   "Senile macular disorder screen", "Digital imaging of retina", 
                                   "[SO]Retina", "Retinal scan - laser", "Macula", "Retina",
                                   "Retinitis", "Fundus flavimaculatus", "Macula observation",
                                    "Superior temporal branch of retinal vein", "Senile macular disorder screening",
                                   "Retinal vein", "Branch of Retinal vein","Retinal artery","Limiting membrane of retina",
                                  "Nerve fibre layer of retina","Inferior temporal quadrant of retina","Central retinal vein",
                                  "Secondary syphilitic chorioretinitis", "Maculopapular", "Parafoveal retina", "Retinal structure",
                                   "Tomaculous neuropathy", "CMV retinitis", "Choroidal and retinal structures", "Retinal arteriole")
```

Exclusion keywords for case
```{r}
exclusion_keyword_patterns_case <- "non-diab|non diab"
```

### Define codes

Inclusion codes
```{r}
inclusion_code_patterns <- "^F42"
inclusion_codes <- c("F4407")
```

Exclusion codes
```{r}
exclusion_code_patterns <- c("^hg1|^hh51|^j24|^kaA|^m[5-7]")
```

### Create diabetic eye disease code dictionary

Import full dictionary
```{r}
full_dict <- readRDS("generated_data/full_dict.RDS")
```

Create eye disease code dictionary
```{r}
eye_disease_dict <- full_dict %>% 
  filter(grepl(inclusion_keyword_patterns, term_description, ignore.case = T) |
           grepl(inclusion_code_patterns, code, ignore.case = T) |
           code %in% inclusion_codes) %>%
  filter(!grepl(global_exclusion_keyword_patterns, term_description, ignore.case = T) &
         !grepl(exclusion_keyword_patterns, term_description, ignore.case = T) &
           !grepl(exclusion_code_patterns, code, ignore.case = F) &
         !(term_description %in% exclusion_keywords))
```

Create diabetic eye disease code dictionary
```{r}
dm_eye_disease_dict <- 
  eye_disease_dict %>% 
  filter(grepl(inclusion_keyword_patterns_case, term_description, ignore.case=T) & 
           !grepl(exclusion_keyword_patterns_case, term_description, ignore.case=T))
```

Generate non-diabetic eye disease code dictionary
```{r}
nondm_eye_disease_dict <-
  eye_disease_dict %>% filter(!(code %in% dm_eye_disease_dict$code))
```

Save diabetic eye disease code dictionary and nondiabetic eye disease code dictionary
```{r}
saveRDS(dm_eye_disease_dict,"generated_data/dm_eye_disease_dict.RDS")
saveRDS(nondm_eye_disease_dict,"generated_data/nondm_eye_disease_dict.RDS")
```

## Prepare diabetic kidney disease code dictionary

### Define Keywords

Inclusion keywords
```{r}
inclusion_keyword_patterns <- "chronic kidney|chronic renal|ckd|glomerul|kidney failure|renal failure|uraemi|uremi|nephropath|proteinuri|albuminuri|nephrotic|nephrosis|nephritic|renal manifestation|renal complication|end stage renal|end stage kidney|esrd"
```

Inclusion keywords for case
```{r}
inclusion_keyword_patterns_case <- "stage 3|stage 4|stage 5|end stage|chronic renal failure|chronic kidney failure|chronic uraemia|G3|G4|G5|A3|diabetes|diabetic|persistent proteinur|persistent albuminur|persistent microalbuminur|persistent macroalbuminur|ns - nephrotic syndrome$|^nephrotic syndrome NOS$"

inclusion_keywords_case <- c("Nephrotic syndrome")
```

Global Exclusion Keywords
```{r}
global_exclusion_keyword_patterns <- "normal| absent|inspection|examination|evaluation|referral|^no |^O/E- no |^O/E - no |excluded|not seen|seen by|not examined|did not attend|assessable|offered|^trauma| trauma|malnutrition|screening for|^FH|family history|interpret|under care|refused|unsuitable|at risk|resolved|disease screening|not indicated|letter|excepted|exception|fear of|education|sharing consent|tumour|pregnancy|gestational|diabetic child|complication of care|steriod-induced|lymphoma|cytomegalovirus|rubella|blastoma|histoplasm|toxoplasm|neoplasm|herpes|tubercul|pediculus| observation$|injury|streptococcal"
```

Exclusion keywords
```{r}
exclusion_keyword_patterns <- "nephritic factor|antibody|disease screening|rate testing|predicted stage|acute|induced by|metals|cadmium|lead|mercury|toxic|abortion|pregnancy|gestational|delivery|calculated by|^glomerular filtration rate$|nephropathy screen|invite|incipient|rate using|laboratory study|monitoring administration|Glomerular function test|benign|haemolytic|B12 deficiency|pyonephrosis|analgesic|Exercise|Adrenal|test strip|proteinuria negative|test urine sample"
```

### Define codes

```{r}
ckd_codelist <- fread("raw_data/kd_code_lists/opensafely-kidney-transplant-2020-07-15.csv") %>%
  full_join(fread("raw_data/kd_code_lists/opensafely-dialysis-2020-07-16.csv")) %>%
  full_join(fread("raw_data/kd_code_lists/opensafely-chronic-kidney-disease-2020-04-14.csv")) %>%
  dplyr::rename(term_description = CTV3PreferredTermDesc) %>%
  dplyr::rename(code = CTV3ID) 
inclusion_codes <- ckd_codelist$code
```

### Create diabetic kidney disease code dictionary

Import full dictionary
```{r}
full_dict <- readRDS("generated_data/full_dict.RDS")
```

Create kidney disease dictionary
```{r}
kidney_disease_dict <- full_dict %>% 
  filter(grepl(inclusion_keyword_patterns,term_description, ignore.case = T) |
           code %in% inclusion_codes) %>%
  filter(!grepl(global_exclusion_keyword_patterns, term_description, ignore.case = T) &
           !grepl(exclusion_keyword_patterns, term_description, ignore.case = T))
```

Generate kidney disease dictionary that will be used for identifying diabetic kidney disease cases
```{r}
kidney_disease_case_dict <- kidney_disease_dict %>% 
  filter(grepl(inclusion_keyword_patterns_case,term_description,ignore.case = T)|
           term_description %in% inclusion_keywords_case)
```

Generate kidney disease dictionary that will be used for identifying controls to be excluded
```{r}
kidney_disease_control_exclusion_dict <- 
  kidney_disease_dict %>% filter(!(code %in% kidney_disease_case_dict$code))
```

Save kidney disease case dictionary and kidney disease control exclusion code dictionary
```{r}
saveRDS(kidney_disease_case_dict,"generated_data/kidney_disease_case_dict.RDS")
saveRDS(kidney_disease_control_exclusion_dict,"generated_data/kidney_disease_control_exclusion_dict.RDS")
```





