# Create biomarker trajectory data from UK biobank assessment center data{#traj-ukb}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=T)
```

```{r, message=F}
library(tidyverse)
library(lubridate)
source("functions.R")
```

Load reformatted UKB demographics and labs tables
```{r}
demog_ukb <- readRDS("generated_data/demog_UKB.RDS")
labs_ukb <- readRDS("generated_data/labs_UKB.RDS")
demog_sel <- readRDS("generated_data/demog_selected.RDS")
```

Define fieldID key
```{r}
fieldID_key <- c(
  "30080" = "platelets",
  "30500" = "albumin_urine",
  "30510" = "creat_urine",
  "30600" = "albumin_blood",
  "30690" = "chol",
  "30700" = "creat_blood",
  "30720" = "cystatin_c",
  "30740" = "glucose_rand", #"Blood glucose"
  "30750" = "hba1c",
  "30760" = "hdl",
  "30780" = "ldl",
  "30870" = "trig",
  "4080" =  "sbp",
  "93" = "sbp",
  "4079" = "dbp",
  "94" = "dbp"
)
```

Create biomarker trajectory data for all biomarkers defined above
```{r}
biomarker_traj_tab_long <- 
  full_join(labs_ukb,demog_ukb) %>% 
  select(f.eid,matches(paste0(paste(paste0('f.',names(fieldID_key)),collapse = '|')))) %>%
  pivot_longer(cols = -f.eid, names_to = "fields", values_to = "measurement") %>%
  filter(!is.na(measurement)) %>%
  separate(fields, c('f',"biomarker_field_id", "visit", "array"), sep = "\\.") %>%
  mutate(biomarker = fieldID_key[biomarker_field_id]) %>% 
  left_join(demog_sel %>% select(f.eid,date_init,date_repeat)) %>%
  mutate(event_dt = as.Date(ifelse(visit == '0', date_init, 
                                   ifelse(visit=='1', date_repeat, NA)), 
                            origin=origin)) %>%
  select(f.eid,measurement,event_dt,visit,biomarker)
```

Compute urine albumin to urine creatinine ratio (UACR)
```{r}
uacr_traj <- biomarker_traj_tab_long %>% filter(biomarker %in% c("albumin_urine","creat_urine")) %>%
  pivot_wider(id_cols = c(f.eid,event_dt), names_from = c(biomarker,visit), values_from = measurement) %>%
  mutate(uacr0 = albumin_urine_0/(creat_urine_0*10^-3), uacr1 = albumin_urine_1/(creat_urine_1*10^-3)) %>%
  select(f.eid,event_dt,uacr0,uacr1) %>%
  pivot_longer(cols = c(uacr0,uacr1), names_to = "visit", values_to = "measurement") %>%
  mutate(visit = ifelse(visit == "uacr0",'0','1')) %>%
  select(f.eid,measurement,event_dt,visit) %>%
  filter(!is.na(measurement)) %>%
  mutate(biomarker = "uacr")
```

Add UACR trajectory data to the `biomarker_traj_tab_long`
```{r}
biomarker_traj_tab_long <- bind_rows(biomarker_traj_tab_long,uacr_traj) %>% 
  filter(!is.na(event_dt),!is.na(measurement))
```

Create macro/microalbuminuria event table
```{r}
albuminuria_event_tab <- 
  biomarker_traj_tab_long %>% 
  filter(biomarker == "uacr") %>%
    mutate(macroalbuminuria = ifelse(measurement < 33.9,F,T),
           microalbuminuria = ifelse(measurement < 3.4,F,T)) %>%
    select(f.eid,event_dt,macroalbuminuria,microalbuminuria) %>%
    pivot_longer(cols=c(macroalbuminuria,microalbuminuria),values_to = "event", names_to = "type") %>%
  select(f.eid,event,event_dt,type)
```

Infer macro/microalbuminuria status:\par

If a subject was missing urine albumin or urine creatinine levels, we cannot compute urine albumin to creatinine ratio (UACR). Urine albumin levels were assumed to be undetectable (NA values) if urine albumin levels were below 6.7. Subjects with known detection issue had missing urine albumin measurements. However, we can still rule out micro/macroalbuminuria using the urine creatinine levels by computing the upper bound of UACR.

To elaborate, suppose a subject's urine albumin level $a$ is below the detection level, i.e. $a < 6.7$. Then, $ACR = \frac{a}{c \times 10^{-3}} < \frac{6.7}{c\times 10^{-3}}$ where $c:=$ urine creatinine levels. Thus, $\frac{6.7}{c\times 10^{-3}}$ is an upper bound of ACR for the subject. Thus

  -if the upper bound of $ACR$ is less than 33.9, rule out macroalbuminuria

  -if the upper bound of $ACR$ is less than 3.4, rule out microalbuminuria.

```{r}
subjects_with_urine_albumin_below_detection_limit <- 
  labs_ukb %>% 
  rename(below_detection_visit0 = f.30505.0.0, below_detection_visit1 = f.30505.1.0) %>%
  select(f.eid,contains("below")) %>%
  pivot_longer(contains("below"),names_to = "visit",values_to = "below_detection_limit") %>%
  mutate(visit = ifelse(visit == "below_detection_visit0",0,1),
         below_detection_limit = ifelse(is.na(below_detection_limit),NA,T)) %>%
  filter(below_detection_limit == T) %>% 
  select(f.eid,visit)

albuminuria_event_tab_inferred <- 
  biomarker_traj_tab_long %>% filter(biomarker == "creat_urine") %>%
  filter(f.eid %in% subjects_with_urine_albumin_below_detection_limit$f.eid) %>%
  mutate(ACR_upper_bound = 6.7/(measurement*10^-3)) %>%
  mutate(macroalbuminuria = ifelse(ACR_upper_bound < 33.9,F,NA),
         microalbuminuria = ifelse(ACR_upper_bound < 3.4,F,NA)) %>% 
  select(f.eid,event_dt,macroalbuminuria,microalbuminuria) %>%
  pivot_longer(cols=c(macroalbuminuria,microalbuminuria),values_to = "event", names_to = "type") %>%
  filter(!is.na(event)) %>%
  select(f.eid,event,event_dt,type)
```

Combine inferred albuminuria event table with `albuminuria_event_tab`
```{r}
albuminuria_event_tab <- bind_rows(albuminuria_event_tab,albuminuria_event_tab_inferred) %>% distinct()
```

Load demographics table and subjects with sex mismatch
```{r}
demog <- readRDS("generated_data/demog_selected.RDS")
sex_mismatch_subject_ids <- readRDS("generated_data/sex_mismatch_subject_ids.RDS")
```

Filter out from from biomarker trajectory table and albuminuria event table:

- any subjects with sex mismatch
- any events with special dates
- any events with unknown date
- any unknown measurements (only applies to biomarker trajectory table)

```{r}
biomarker_traj_tab_long <- biomarker_traj_tab_long[!(biomarker_traj_tab_long$f.eid %in% sex_mismatch_subject_ids),]
biomarker_traj_tab_long <- biomarker_traj_tab_long %>% left_join(demog %>% select(f.eid,DOB),by="f.eid")
biomarker_traj_tab_long$event_dt <- cleandates(biomarker_traj_tab_long$event_dt,biomarker_traj_tab_long$DOB,T) %>% as.Date()
biomarker_traj_tab_long <- biomarker_traj_tab_long %>% select(-DOB)
biomarker_traj_tab_long <- biomarker_traj_tab_long[!is.na(biomarker_traj_tab_long$event_dt),]
biomarker_traj_tab_long <- biomarker_traj_tab_long[!is.na(biomarker_traj_tab_long$measurement),]

albuminuria_event_tab <- albuminuria_event_tab[!(albuminuria_event_tab$f.eid %in% sex_mismatch_subject_ids),]
albuminuria_event_tab <- albuminuria_event_tab %>% left_join(demog %>% select(f.eid,DOB),by="f.eid")
albuminuria_event_tab$event_dt <- cleandates(albuminuria_event_tab$event_dt,albuminuria_event_tab$DOB,T) %>% as.Date()
albuminuria_event_tab <- albuminuria_event_tab %>% select(-DOB)
albuminuria_event_tab <- albuminuria_event_tab[!is.na(albuminuria_event_tab$event_dt),]
```

Save the biomarker trajectory table and albuminuria event table extracted from the UKB asssessment center data.
```{r}
saveRDS(biomarker_traj_tab_long,"generated_data/biomarker_trajectory_ukb.RDS")
saveRDS(albuminuria_event_tab,"generated_data/albuminuria_event_tab_ukb.RDS")
```




