# Create biomarker trajectory data from primary care data {#traj-pcp}

In this chapter, we create biomarker trajectory data using the primary care data. The biomarkers we phenotype are

- urine albumin 
- urine creatinine
- blood creatinine
- blood glucose (random)
- fasting glucose
- hba1c
- HDL
- LDL
- triglycerides
- systolic blood pressure (SBP)
- diastolic blood pressure (DBP)
- urine albumin to urine creatinine ratio (UACR)
- BMI

As we did in the previous chapter, we phenotype patient's macroalbumuminuria or microalbuminuria status, and exclude events whose measurements or their dates are missing. However, we do not apply sex mismatch filters since we have already filtered out those subjects in chapter \@ref(prep-pcp-data).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=F)
```

Load necessary packages.
```{r, message=F, warning=F}
library(tidyverse)
library(knitr)
```

Load reformatted primary care table.
<!-- and remove events marked as special dates -->
<!-- # gp_clinical <- gp_clinical[!gp_clinical$special_dt,] -->
```{r, message=F}
gp_clinical <- data.table::fread("generated_data/entire_gp_clinical_30March2021_formatted.txt")
head(gp_clinical)
```

Read in the previously generated full code dictionary for primary care terms. 
```{r}
full_dict <- readRDS("generated_data/full_dict.RDS")
tail(full_dict)
```

For the purpose of combining the dictionary with gp_clinical, remove terms with missing descriptions, additional descriptions for the same code, and identical codes+descriptions across terminology(read v2/CTV3).
```{r, message=F, warning=F}
full_dict <- full_dict %>%
  select(-terminology) %>%
  distinct() %>%
  filter(term_description != "") %>%
  distinct(code, .keep_all = T)

tail(full_dict)
```
Add the term descriptions to gp_clinical.
```{r, message=F}
gp_clinical <- gp_clinical %>%
  left_join(full_dict) 
```

Define the clinical terms to extract from gp_clinical for each biomarker. Each biomarker gets a string to be used in a grepl pattern matching query, and as such different patterns should be separated by '|', brackets used to denote multiple possible patterns to match, and '^' to denote the beginning of the string.
```{r}
BP_codes <- "^246[.cdgABCDEFGJNPQRSTVWXY012345679]|^XaF4[abFLOS]|^XaJ2[EFGH]|^XaKF[xw]|^G20"
HDL_codes <- '^44d[23]|^X772M|^44P[5BC]|^XaEVr'
LDL_codes <- '^44d[45]|^44P[6DE]|^XaEVs' 
totchol_codes <- "^44P[.12349HJKZ]|^XE2eD|^XSK14|^XaFs9|^XaIRd|^XaJe9|^XaLux" 
triglyc_codes <- '^44e|^44Q|^X772O|^XE2q9' 
fastgluc_codes <- "^44[fg]1"
randgluc_codes <- "^44[fg][0\\.]|^44TA|^XM0ly"
a1c_codes <- "^XaPbt|^XaERp|^X772q|^42W[12345Z\\.]\\.|^44TB\\."
height_weight_BMI_codes <- "^XaCDR|^XaJJH|^XaJqk|^XaZcl|^22K|^229|^22A|^162[23]|^X76CG|^XE1h4|^XM01G|^Xa7wI"
blood_creatinine_codes <- '^44J3[.0123z]|^44J[CDF]|XE2q5|XaERc|XaERX|XaETQ|^4Q40.|X771Q'
urine_creatinine_codes <- '^46M7'
urine_albumin_codes <- '^46N4|^XE2eI|^46N8.|^46W[\\.01]|^XE2bw'
UACR_codes <- '^46TC|^XE2n3|^X773Y|^46TD|^XE2n4'
```

## Total cholesterol
UK Biobank measurements range from 0.601 to 15.46 mmol/L.  We filter the allowable range to lie between 0.5 and 30.
```{r, message=F, warning=F}
totchol <- gp_clinical %>%
  filter(grepl(totchol_codes, code)) %>%
  mutate(totchol = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(totchol > 0.5 & totchol < 30) %>%
  filter(!(value3 %in% c("MEA001", "mg/mmol", "%", "MEA090", "MEA093", 
                         "MEA099", "MEA110", "MEA142",  "MEA156", "mmo", 
                         "mmol", "mmol/d", "mOsm/kg", "mPa/s", "nmol/l", 
                         "U/mL", "umol/L")))

totchol %>% group_by(code, term_description) %>% 
  summarize(n=n(), mean = mean(totchol, na.rm=T)) %>% 
  arrange(desc(n)) %>% kable()

ggplot(data=totchol, aes(x=log10(totchol))) + geom_density() + theme_minimal()
```

## HDL
UK Biobank measurements range from 0.219 to 4.401 mmol/L.  We filter the allowable range to be under 10, but not 0.
```{r, warning=F, message=F}
HDL <- gp_clinical %>%
  filter(grepl(HDL_codes, code)) %>% 
  mutate(HDL = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(HDL > 0 & HDL < 10) %>%
  filter(!(value3 %in% c("09", "MEA082", "MEA095", "MEA099", "MEA104", "MEA151", 
                         "mmol", "mOsm/kg", "mg/mmol")))

HDL %>% group_by(code, term_description) %>% 
  summarize(n=n(), mean = mean(HDL, na.rm=T)) %>% 
  arrange(desc(n)) %>% kable()

ggplot(data=HDL, aes(x=log10(HDL))) + geom_density() + theme_minimal()
```

## LDL
UK Biobank measurements range from 0.226 to 9.797 mmol/L.  We filter the allowable range to be under 30, but not 0.
```{r, warning=F, message=F}
LDL <- gp_clinical %>%
  filter(grepl(LDL_codes, code)) %>%
  mutate(LDL = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(LDL > 0 & LDL < 30) %>%
  filter(!(value3 %in% c("MEA142", "MEA151")))

LDL %>% group_by(code, term_description) %>% 
  summarize(n=n(), mean = mean(LDL, na.rm=T)) %>% 
  arrange(desc(n)) %>% kable()

ggplot(data=LDL, aes(x=log10(LDL))) + geom_density() + theme_minimal()
```

## Triglycerides
UK Biobank measurements range from 0.231 to 11.278 mmol/L. We filter the allowable range to be under 30, but not 0.
```{r, warning=F, message=F}
triglycerides <- gp_clinical %>%
  filter(grepl(triglyc_codes, code)) %>%
  mutate(triglyc = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(triglyc > 0 & triglyc < 30) %>%
  filter(!(value3 %in% c("g", "g/l", "MEA099", "MEA156")))

triglycerides %>% group_by(code, term_description) %>% 
  summarize(n=n(), mean= mean(triglyc, na.rm=T)) %>% 
  arrange(desc(n)) %>% kable()

ggplot(data=triglycerides, aes(x=log10(triglyc))) + geom_density() + theme_minimal()
```

## Fasting Glucose
UK Biobank measurements for glucose range from 0.995 to 36.813 mmol/L.  We filter the allowable range of fasting glucose to be under 60, but not 0. 
```{r, message=F, warning=F}
fastgluc <- gp_clinical %>%
  filter(grepl(fastgluc_codes, code)) %>%
  mutate(fastgluc = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(fastgluc > 0 & fastgluc < 60) %>%
  filter(value3 %in% c("", "MEA000", "MEA061", "MEA096", "MEA194", "Mmol/L", 
                       "Unknown", "mmol/L", "mmol/l", "mU/l", "units"))

fastgluc %>% group_by(code, term_description) %>% 
  summarize(n=n(), mean = mean(fastgluc, na.rm=T)) %>% 
  arrange(desc(n)) %>% kable()

ggplot(data=fastgluc, aes(x=log10(fastgluc))) + geom_density() + theme_minimal()
```

## Random glucose
UK Biobank measurements for glucose range from 0.995 to 36.813 mmol/L.  We filter the allowable range of fasting glucose to be under 60, but not 0.
```{r, message=F, warning=F}
randgluc <- gp_clinical %>%
  filter(grepl(randgluc_codes, code)) %>% 
  mutate(randgluc = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(randgluc > 0 & randgluc < 60) %>%
  filter(value3 %in% c("", "MEA000", "MEA061", "MEA096", "MEA194", "Mmol/L", 
                       "Unknown", "mmol/L", "mmol/l", "mU/l", "units"))

randgluc %>% group_by(code, term_description) %>% 
  summarize(n=n(), mean = mean(randgluc, na.rm=T)) %>% 
  arrange(desc(n)) %>% kable()

ggplot(data = randgluc, aes(x=log10(randgluc))) + geom_density() + theme_minimal()
```

## Urine Albumin
UK Biobank measurements for microalbumin in urine have a maximum value of 6746.5 mg/L, with many values below the measurement instrument's lower limit of detection of 6.7 mg/L. We filter the allowable range to be under 7,000, but not 0.
```{r, message=F, warning=F}
#Extract urine albumin
ualb <- gp_clinical %>%
  filter(grepl(urine_albumin_codes, code, ignore.case=T)) %>%
  mutate(micro=grepl("micro", term_description, ignore.case=T)) %>%
  mutate(value = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(!is.na(value)  & value != 0) %>% #& value < 500
  filter(!(value3 %in% c("0.3", "MEA001", "mg/mmol", "ml/min", "mmol/L", "MEA086",
                         "MEA096", "MEA120", "MEA142", "MEA166", "g/L", "MEA057"))) %>%  
  filter(value < 7000) %>%
  select(-value1, -value2) %>% 
  dplyr::rename(code_ualb = code, term_description_ualb = term_description,
                             ualb_mg_L = value) %>%
  distinct()

ualb %>% group_by(code_ualb, term_description_ualb) %>%
  summarize(n=n(), mean = mean(ualb_mg_L)) %>%
  arrange(desc(n)) %>% kable()

```

Based on the distributions, albumin values are given in mg/L, regardless of whether the term description says microalbumin or albumin. 
```{r, message=F, warning=F}
ggplot(data=ualb, aes(x=log10(ualb_mg_L), color = term_description_ualb)) + geom_density()
```

## Urine creatinine
UK Biobank measurements range from 88 to 88,160 micromoles/L. We filter the allowable range to be below 88,000, but not 0.
```{r, message=F, warning=F}
ucreat <- gp_clinical %>%
  filter(grepl(urine_creatinine_codes, code, ignore.case=T)) %>%
  mutate(value = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(!is.na(value)  & value != 0) %>% 
  mutate(ucreat_umol_L = value * 1000) %>%
  filter(value3 != "MEA099" & value3 != "umol/l") %>%
  filter(ucreat_umol_L < 88000) %>%
  select(-value, -value1, -value2, -value3) %>%
  dplyr::rename(code_ucreat = code, term_description_ucreat = term_description) %>%
  distinct()

ucreat %>% group_by(code_ucreat, term_description_ucreat) %>%
  summarize(n=n(), mean = mean(ucreat_umol_L)) %>% kable()

ggplot(data=ucreat, aes(x=log10(log10(ucreat_umol_L)))) + geom_density() + theme_minimal()
```

## Urine ACR (Albumin:Creatinine Ratio)
This ratio will be needed to define micro/macroalbuminuria phenotypes. The calculated UACR values in UK Biobank lie between 0.15 and 2141.17.  For UACR calculated from primary care albumin and creatinine, we only filter based on the original albumin and creatinine measurements. For primary care records of UACR, we filter the value to be below 1000, but not 0.
```{r, message=F, warning=F}
#Manually generate the UACRs from urine creatinine and albumin.
ratio_manual <- ualb %>%
  inner_join(ucreat) %>%
  mutate(UACR = ualb_mg_L/ucreat_umol_L *1000, manual=T)

#Extract UACR from primary care that is already calculated
ratio <- gp_clinical %>%
  filter(grepl(UACR_codes, code, ignore.case=T)) %>%
  mutate(UACR = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3)),
         micro = grepl("micro", term_description),
         manual = F) %>%
  filter(!is.na(UACR)  & UACR != 0 & UACR < 1000) %>%
  filter(!(value3 %in% c("%", "MEA001", "MEA083", "MEA096",
                         "MEA156", "MEA169", "MEA241",
                         "ml/min", "mg/L", "m1/min", "mmol/L", 
                         "ng/mL","umol/L", "mg/l"))) %>%
  select(-value1, -value2, -value3) %>%
  distinct() %>%
  dplyr::rename(code_uacr = code, term_description_uacr = term_description)

#Combine the manually-calculated and already-calculated values of UACR
UACR <- full_join(ratio, ratio_manual) %>%
  distinct(f.eid, event_dt, .keep_all = T)

UACR %>% group_by(manual, term_description_uacr, term_description_ualb) %>%
  summarize(n=n(), mean = mean(UACR, na.rm=T)) %>%
  arrange(desc(n)) %>% kable()

ggplot(data=UACR, aes(x=log10(UACR))) + geom_density() + theme_minimal()
```

## Blood/Serum creatinine
This is needed to calculate eGFR. UK Biobank measures lie between 10.7 and 1499.3. We filter the allowable range to lie between 10 and 1500.
```{r, warning=F, message=F}
blood_creatinine <- gp_clinical %>%
  filter(grepl(blood_creatinine_codes, code)) %>% 
  mutate(blood_creatinine = coalesce(as.numeric(value1), as.numeric(value2))) %>%
  filter(blood_creatinine > 10 & blood_creatinine < 1500) %>%
  filter(!(value3 %in% 
             c("m1/min", "MEA082", "MEA083", "MEA095", "MEA096", "mg/mmol", 
               "ml/min", "mmol", "mmol/l", "mmol/L", "nmol/l", "pmol/l", 
               "uL/cu mm", "um", "umo", "umol"))) %>% 
  mutate(creatinine_type = ifelse(grepl("serum", term_description, ignore.case=T), "Serum", 
                               ifelse(grepl("plasma", term_description, ignore.case=T), "Plasma",
                                      "Unspecified"))) 

blood_creatinine %>% group_by(code, term_description) %>% 
  summarize(n=n(), mean=mean(blood_creatinine)) %>% 
  arrange(desc(n)) %>% kable()

ggplot(data=blood_creatinine, aes(x=blood_creatinine, color=creatinine_type)) +
  geom_density() + theme_minimal()
```

## HbA1c
Calculate HbA1c in both % (US) and mmol/mol (UK). UK Biobank measurements lie between 15 and 515.2 mmol/mol.  In primary care data, some values are given in percent, some in mmol/mol. We filter the allowable range to lie between 4 and 18 (%) or between 20.2 and 173.2 (mmol/mol) to guard against including values recorded under the incorrect unit. Conversion between % and mmol/mol is performed using the IFCC-NGSP master equation (https://doi.org/10.1373/clinchem.2008.103556).

```{r, message=F, warning=F}
a1c <- gp_clinical %>%
  filter(grepl(a1c_codes, code)) %>%
  mutate(hba1c = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(hba1c > 0) %>%
  mutate(value3 = toupper(value3)) %>%
  mutate(value3 = ifelse(value3 %in%  c("MEA000", "MMOL/M", "MEA097", "UNKNOWN", "MEA001", 
                                        "%", "HBA1C", "%TOTAL HB", "% TOTAL HB", "MEA215", 
                                        "MMOL/MOL HB", "PER CENT", "%TOTAL"), "", value3)) %>%
  mutate(units = ifelse(value3 != "", value3, 
                        ifelse(code %in% c("XaPbt", "42W5."), "MMOL/MOL", "%"))) %>%
  filter(units %in% c("%", "MMOL/MOL")) %>%
  mutate(hba1c_percent = ifelse(units == "%", round(hba1c, 1), 
                                round(hba1c/10.929 + 2.15, 1))) %>%
  mutate(hba1c_mmol_mol = ifelse(units =="%", round(10.929 * (hba1c - 2.15), 1), 
                                 round(hba1c, 1))) %>%
  filter(hba1c_percent > 4 & hba1c_percent < 18) %>%
  dplyr::rename(original_unit = units)

a1c %>% group_by(code, term_description) %>%
  summarize(n=n(), mean=mean(hba1c_percent, na.rm=T)) %>%
  arrange(desc(n)) %>% kable()

ggplot(data=a1c, aes(x = hba1c_percent, color = original_unit)) + geom_density()
```

## BMI
Height may be measured using meters or centimeters. Height was filtered to lie between 1.25m (4ft 1.2 inches) and 2.1m (6ft 10.7 inches). Values between 125 and 210 were assumed to be in cm and so were divided by 100.
Weight is measured in kg and was filtered to lie between 30kg(66.1 lb) and 200kg(440.9 lb).
BMI was filtered to lie between 12 and 75 kg/m^2. 
```{r, message=F, warning=F}
#Extract height, weight, and BMI
hwbmi <- gp_clinical %>%
  filter(grepl(height_weight_BMI_codes, code)) %>%
  mutate(value = coalesce(as.numeric(value1), as.numeric(value2), as.numeric(value3))) %>%
  filter(!is.na(value) & value > 0) %>%
  mutate(trait = ifelse(grepl("BMI|Body Mass Index", term_description, ignore.case=T), "BMI",
                        ifelse(grepl("Height", term_description, ignore.case=T), "Height",
                               "Weight"))) %>%
  mutate(value = ifelse(trait == "Height" & value > 2.1, value/100, value)) %>% #cm to meters
  filter((trait == "Weight" & value < 200 & value > 30) |
         (trait == "Height" & value < 2.1 & value > 1.25 )|
         (trait == "BMI" & value < 75 & value > 12)) %>% 
  arrange(f.eid, event_dt)

#Here are the resulting terms:
hwbmi %>% group_by(code, term_description, trait) %>% 
  summarize(n = n(), mean=round(mean(value), 1)) %>% 
  arrange(trait, desc(n)) %>% kable()

#Separate the traits
weight <- hwbmi %>% filter(trait == "Weight") 
height <- hwbmi %>% filter(trait == "Height") 
BMI <- hwbmi %>% filter(trait == "BMI") 

#Some records from data provider 2 give weight and BMI
bmi_extra <- weight %>% 
  filter(data_provider == 2) %>% 
  mutate(bmi_extra = as.numeric(value3)) %>% 
  filter(!is.na(bmi_extra)) %>%
  filter(bmi_extra > 12)
head(bmi_extra)

#combine the traits in wider format
weight_pre <- weight %>% select(-c(terminology, value1, value2, value3, trait)) %>%
  dplyr::rename(weight = value, weight_code = code, weight_term_description = term_description)
height_pre <- height %>% select(-c(terminology, value1, value2, value3, trait))  %>%
  dplyr::rename(height = value, height_code = code, height_term_description = term_description)
BMI_pre <- BMI %>% select(-c(terminology, value1, value2, value3, trait))  %>%
  dplyr::rename(BMI = value, BMI_code = code, BMI_term_description = term_description)
bmi_extra_pre <- bmi_extra %>% select(-c(terminology, term_description, value1, value2, value3, trait, value)) %>%
  dplyr::rename(bmi_extra_code = code)

joined_bmi <- full_join(weight_pre, height_pre) %>%
              full_join(BMI_pre) %>%
              full_join(bmi_extra_pre)
head(joined_bmi)
```

Clean the combined data, filling in missing height values from previous/subsequent measurements or BMI. Filter out cases where the calculated and reported BMI differ by more than 1.5.
```{r, message=F, warning=F}
cleaned_bmi <- joined_bmi %>%
  group_by(f.eid) %>%
  fill(height, .direction = "downup") %>%
  mutate(BMI_coalesce = round(coalesce(BMI, bmi_extra), 1),
         BMI_calculated = round(weight/(height^2), 1),
         BMI_mismatch = BMI_coalesce != BMI_calculated, 
         BMI_diff = BMI_coalesce - BMI_calculated) %>%
  filter(!(!is.na(BMI_diff) & abs(BMI_diff) > 1.5)) %>%
  mutate(BMI_final = coalesce(BMI_calculated, BMI_coalesce)) %>%
  mutate(height = ifelse(!is.na(height), height, sqrt(weight/BMI))) %>%
  dplyr::rename(BMI_reported = BMI_coalesce, height_carried = height) %>%
  filter(height_carried < 2.1 & height_carried > 1.25 & 
           BMI_final < 75 & BMI_final > 12) %>%
  select(f.eid, data_provider, event_dt, weight, height_carried, 
         BMI_reported, BMI_calculated, BMI = BMI_final) %>% 
  distinct() %>%
  mutate(source="PC")

ggplot(data=cleaned_bmi, aes(x=log10(BMI))) + geom_density() + theme_minimal()
```

```{r, include=FALSE}
#bp$term_description[bp$term_description == ""] <- NA
```

## Blood Pressure
To to extract blood pressure, it is necessary to consider both the cases where both Systolic and Diastolic BP are provided in the same record, as well as cases where Systolic and Diastolic BP are given in separate records. 
```{r, message=F, warning=F}
bp <- gp_clinical %>%
  filter(grepl(BP_codes, code)) %>%
  filter(value1 != "" | value2 != "" | value3 != "")  %>%
  mutate(value1 = as.numeric(value1), value2 = as.numeric(value2), value3 = as.numeric(value3)) 

#Remove some small values in value3 of unknown meaning
bp$value3[bp$value3 < 10] <- NA

#Separate the data into two types of record: One where 2 values are given in 1 record (e.g. Systolic and Diastolic) and another where there is only one value (i.e Systolic only or Diastolic only) given per record. 
```

Multiple values per record: Take the larger value to be systolic and the smaller value to be diastolic. Filter out any records where either of these values are 0.
```{r, message=F, warning=F}
bp_mult <- bp %>% 
  rowwise() %>%
  filter(sum(!is.na(value1), !is.na(value2), !is.na(value3)) == 2) %>%
  ungroup() %>%
  mutate(Systolic_bp_pc = pmax(value1, value2, value3, na.rm=T)) %>%
  mutate(Diastolic_bp_pc = pmin(value1, value2, value3, na.rm=T)) %>%
  filter(Systolic_bp_pc != 0 & Diastolic_bp_pc != 0)
```

One value per record: filter out those with a value of 0. 
```{r, message=F, warning=F}
bp_single <- bp %>% 
  rowwise() %>%
  filter(sum(!is.na(value1), !is.na(value2), !is.na(value3)) == 1) %>%
  ungroup() %>%
  mutate(value = coalesce(value1, value2, value3)) %>%
  filter(value != 0) %>%
  arrange(f.eid, event_dt) %>%
  mutate(bp_type =  ifelse(grepl("systolic", term_description, ignore.case=T), "Systolic_bp_pc",
               ifelse(grepl("diastolic", term_description, ignore.case = T), "Diastolic_bp_pc", "Unknown"))) %>%
  select(-value1, -value2, -value3) %>%
  distinct()

#Here are the results for single type:
bp_single %>% 
  group_by(code, term_description, bp_type) %>% 
  summarize(n=n(), mean=round(mean(value, 1))) %>% 
  arrange(bp_type, desc(n)) 
```

Look at the remaining codes to see if they are systolic or diastolic.  For many of these, the same code is given twice, each with a different value. Sometimes an record is a duplicate of a systolic or diastolic measurement. If there are two *unique* values given per ID/date, then we can assume they are systolic (higher) and diastolic (lower). Otherwise, discard that set of values. 
```{r, message=F, warning=F}
#Unknowns - not specified as Diastolic vs. Systolic
unknowns <- bp_single %>%
  group_by(f.eid, event_dt) %>%
  filter(sum(bp_type == "Unknown") > 0) %>%
  mutate(n = length(unique(value))) %>%
  filter(n == 2) %>%
  mutate(Systolic_bp_pc = max(value)) %>%
  mutate(Diastolic_bp_pc = min(value)) %>%
  distinct(f.eid, event_dt, value, .keep_all = T) %>%
  group_by(f.eid, event_dt) %>%
  mutate(code_systolic = code[which(value == max(value))], 
         code_diastolic = code[which(value == min(value))], 
         term_description_systolic = term_description[which(value == max(value))], 
         term_description_diastolic = term_description[which(value == min(value))])

#Prepare for merging
bp_single_less <- bp_single %>% 
  distinct(f.eid, data_provider, event_dt, value, bp_type, .keep_all = T) %>%
  group_by(f.eid, event_dt) %>%
  filter(sum(bp_type == "Systolic_bp_pc") == 1 & 
           sum(bp_type == "Diastolic_bp_pc") == 1) %>%
  filter(bp_type != "Unknown")

systolic <- bp_single_less %>%
  filter(bp_type == "Systolic_bp_pc") %>%
  dplyr::rename(code_systolic = code, 
                Systolic_bp_pc = value, 
                term_description_systolic = term_description) %>%
  select(-bp_type)

diastolic <- bp_single_less %>%
  filter(bp_type == "Diastolic_bp_pc") %>%
  dplyr::rename(code_diastolic = code, 
                Diastolic_bp_pc = value, 
                term_description_diastolic = term_description) %>%
  select(-bp_type)
        
bp_single_wide <- full_join(systolic, diastolic)
head(bp_single_wide)
```

Combine each of the cleaned subsets and implement some common sense filters (45 < Systolic bp < 300, 30 < Diastolic bp < Systolic bp).
```{r, messsage=F, warning=F}
unknowns_less <- unknowns %>% 
  select(-term_description, -value, -bp_type, -code, -n) %>%
  distinct()

bp_mult_less <- bp_mult %>% 
  dplyr::rename(term_description_both = term_description) %>%
  dplyr::rename(code_both = code) %>%
  select(f.eid, event_dt, data_provider, terminology, Systolic_bp_pc, Diastolic_bp_pc, code_both, term_description_both) %>%
  distinct()

full_bp_clean <- full_join(unknowns_less, bp_single_wide) %>%
  full_join(bp_mult_less) %>%
  filter(Systolic_bp_pc > Diastolic_bp_pc) %>%
  filter(Systolic_bp_pc >= 45 & Systolic_bp_pc <= 300) %>%
  filter(Diastolic_bp_pc >= 30) %>%
  distinct(f.eid, event_dt, Systolic_bp_pc, Diastolic_bp_pc, .keep_all=T)

head(full_bp_clean)

ggplot(data=full_bp_clean, aes(x=log10(Systolic_bp_pc))) + geom_density() + theme_minimal()
ggplot(data=full_bp_clean, aes(x=log10(Diastolic_bp_pc))) + geom_density() + theme_minimal()
```

## Reformat and trim trajectory table for each biomarker

```{r}
sbp <- full_bp_clean %>% 
  select(f.eid,Systolic_bp_pc,event_dt) %>% 
  rename(measurement = Systolic_bp_pc) %>% 
  mutate(biomarker = "sbp")

dbp <- full_bp_clean %>% 
  select(f.eid,Diastolic_bp_pc,event_dt) %>% 
  rename(measurement = Diastolic_bp_pc) %>% 
  mutate(biomarker = "dbp")

cholesterol <- totchol %>% 
  select(f.eid,totchol,event_dt) %>% 
  rename(measurement = totchol) %>% 
  mutate(biomarker = "chol")

hdl <- HDL %>% 
  select(f.eid,HDL,event_dt) %>% 
  rename(measurement = HDL) %>% 
  mutate(biomarker = "hdl")

ldl <- LDL %>% 
  select(f.eid,LDL,event_dt) %>% 
  rename(measurement = LDL) %>% 
  mutate(biomarker = "ldl")

trigly <- triglycerides %>% 
  select(f.eid,triglyc,event_dt) %>% 
  rename(measurement = triglyc) %>% 
  mutate(biomarker = "trig")

glu_fast <- fastgluc %>%
  select(f.eid,fastgluc,event_dt) %>% 
  rename(measurement = fastgluc) %>%
  mutate(biomarker = "glucose_fast")

glu_rand <- randgluc %>%
  select(f.eid,randgluc,event_dt) %>% 
  rename(measurement = randgluc) %>% 
  mutate(biomarker = "glucose_rand")

hba1c <- a1c %>% 
  select(f.eid,hba1c_mmol_mol,event_dt) %>% 
  rename(measurement = hba1c_mmol_mol) %>%
  mutate(biomarker = "hba1c")

bmi <- cleaned_bmi %>%
  select(f.eid,BMI,event_dt) %>%
  rename(measurement = BMI) %>%
  mutate(biomarker = "bmi")

blood_creat <- blood_creatinine %>%
  select(f.eid,blood_creatinine,event_dt) %>%
  rename(measurement = blood_creatinine) %>%
  mutate(biomarker = "creat_blood")

urine_creat <- ucreat %>% 
  select(f.eid, ucreat_umol_L, event_dt) %>%
  rename(measurement = ucreat_umol_L) %>%
  mutate(biomarker = "creat_urine")

urine_albumin <- ualb %>%
  select(f.eid, ualb_mg_L, event_dt) %>%
  rename(measurement = ualb_mg_L) %>%
  mutate(biomarker = "albumin_urine")

uacr <- UACR %>%
  select(f.eid,UACR,event_dt) %>%
  rename(measurement = UACR) %>%
  mutate(biomarker = "uacr")
```

Combine all of biomarker trajectory datasets into a long format. 
```{r}
biomarker_traj_tab_long <- bind_rows(list(sbp,dbp,cholesterol,hdl,ldl,trigly,
                                          glu_fast,glu_rand,hba1c,bmi,blood_creat,
                                          urine_creat,urine_albumin,uacr))
```

Create macro/microalbuminuria event table.
```{r}
albuminuria_event_tab <- 
  biomarker_traj_tab_long %>% 
  filter(biomarker == "uacr") %>%
    mutate(macroalbuminuria = ifelse(measurement < 33.9,F,T),
           microalbuminuria = ifelse(measurement < 3.4,F,T)) %>%
    select(f.eid,event_dt,macroalbuminuria,microalbuminuria) %>%
    pivot_longer(cols=c(macroalbuminuria,microalbuminuria),values_to = "event", names_to = "type") %>%
  select(f.eid,event,event_dt,type)
```

Filter out the following from the biomarker trajectory table and albuminuria event table:

- any unknown measurements
- any measurements with unknown dates

```{r}
biomarker_traj_tab_long <- biomarker_traj_tab_long[!is.na(biomarker_traj_tab_long$event_dt),]
biomarker_traj_tab_long <- biomarker_traj_tab_long[!is.na(biomarker_traj_tab_long$measurement),]

albuminuria_event_tab <- albuminuria_event_tab[!is.na(albuminuria_event_tab$event_dt),]
```

Save the biomarker trajectory data and albuminuria event table extracted from the primary care data.
```{r}
saveRDS(biomarker_traj_tab_long,"generated_data/biomarker_trajectory_pcp.RDS")
saveRDS(albuminuria_event_tab,"generated_data/albuminuria_event_tab_pcp.RDS")
```










