# Combine biomarker trajectory extracted from from UKB and PCP data {#traj-combine}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, eval = T)
```

```{r}
library(tidyverse)
source("functions.R")
```

Load UKB and PCP biomarker trajectory data
```{r}
biomarker_traj_pcp <- readRDS("generated_data/biomarker_trajectory_pcp.RDS")
biomarker_traj_ukb <- readRDS("generated_data/biomarker_trajectory_ukb.RDS")
```

Combine UKB and PCP biomarker trajectory data
```{r}
biomarker_traj <- bind_rows(biomarker_traj_pcp,biomarker_traj_ukb %>% select(-visit))
biomarker_traj <- biomarker_traj %>% distinct()
```

Load UKB and PCP albuminuria event table
```{r}
albuminuria_event_tab_pcp <- readRDS("generated_data/albuminuria_event_tab_pcp.RDS")
albuminuria_event_tab_ukb <- readRDS("generated_data/albuminuria_event_tab_ukb.RDS")
```

Combine albuminuria event tables
```{r}
albuminuria_event_tab <- bind_rows(albuminuria_event_tab_pcp,albuminuria_event_tab_ukb) %>% distinct()
```

Create micro and microalbuminuria first identification event table
```{r}
microabu_firstoccur <- albuminuria_event_tab %>% 
  filter(type == "microalbuminuria",event == T) %>% select(f.eid,event,event_dt) %>%
  group_by(f.eid) %>% arrange(event_dt) %>% slice(1)
macroabu_firstoccur <- albuminuria_event_tab %>% 
  filter(type == "macroalbuminuria",event == T) %>% select(f.eid,event,event_dt) %>%
  group_by(f.eid) %>% arrange(event_dt) %>% slice(1)
```

Load demographic table
```{r}
demog <- readRDS("generated_data/demog_selected.RDS")
```

Compute eGFR and create eGFR trajectory data
```{r}
creat_blood_traj <- biomarker_traj[biomarker_traj$biomarker == "creat_blood",] %>%
  left_join(demog %>% select(f.eid, SEX, DOB))

creat_blood_traj$age_egfr <-
  lubridate::decimal_date(creat_blood_traj$event_dt) - lubridate::decimal_date(creat_blood_traj$DOB)

creat_blood_traj$measurement_egfr <-
  mapply(compute_egfr,creat_blood_traj$measurement,creat_blood_traj$age_egfr,creat_blood_traj$SEX)

egfr_traj <- creat_blood_traj %>% select(f.eid,measurement_egfr,event_dt) %>%
  rename(measurement = measurement_egfr) %>% mutate(biomarker = "egfr")

egfr_traj <- egfr_traj[!is.na(egfr_traj$measurement),]
```

Define a “prolonged low eGFR event” as as an event with two or more low eGFR event (eGFR < 60) for more than 90 days without a normal eGFR in between (which can be ascertained only when we have at least two low eGFR event 90 days apart). `low_egfr_firstoccur` contains the first occurrence date of the “prolonged low eGFR event.” This event table will be used in phenotyping DKD cases later.
```{r}
prolonged_low_egfr_firstoccur <- egfr_traj %>% 
  mutate(eGFR_lt60 = ifelse(measurement < 60, 1, 0)) %>%
  group_by(f.eid) %>%
  arrange(f.eid, event_dt) %>%
    mutate(switched = eGFR_lt60 != lag(eGFR_lt60)) %>%
    mutate(run_n = cumsum(ifelse(is.na(switched), 0, switched))) %>%
    group_by(f.eid, run_n) %>% 
    mutate(max_days = 
             (lubridate::decimal_date(max(event_dt, na.rm=T)) - 
                lubridate::decimal_date(min(event_dt, na.rm=T)))*365.25) %>%
    filter(eGFR_lt60 & max_days >=90) %>% # for a subject with only one event of egfr < 60, then we would have max_days = 0
    ungroup() %>% 
    group_by(f.eid) %>% arrange(event_dt) %>%
    slice(1) %>% # selecting for the start date of the first period where a subject had egfr level less 60 for 
                                  # 90+ days.
    ungroup() %>% select(f.eid, event_dt)
```

Add eGFR trajectory data to `biomarker_traj`
```{r}
biomarker_traj <- bind_rows(biomarker_traj,egfr_traj)
biomarker_traj <- biomarker_traj[!is.na(biomarker_traj$measurement) & !is.na(biomarker_traj$event_dt),]
```

Break up biomarker trajectory table per biomarker
```{r}
biomarker_traj_list_per_biomarker <- lapply(unique(biomarker_traj$biomarker),function(biomarker){
  biomarker_traj[biomarker_traj$biomarker == biomarker,]
})
names(biomarker_traj_list_per_biomarker) <- unique(biomarker_traj$biomarker)
```

Save biomarker trajectory table
```{r}
saveRDS(biomarker_traj,"generated_data/biomarker_trajectory.RDS")
```

Save trajectory table for each biomarker
```{r}
for(i in 1:length(biomarker_traj_list_per_biomarker)){
    biomarker <- names(biomarker_traj_list_per_biomarker)[i]
    saveRDS(biomarker_traj_list_per_biomarker[[i]],paste0("generated_data/","trajectory_",biomarker,".RDS"))
}
```

Save albuminuria table
```{r}
saveRDS(albuminuria_event_tab,"generated_data/albuminuria_event_tab.RDS")
```

Save macroalbuminuria and microalbuminuria first occurrence table
```{r}
saveRDS(macroabu_firstoccur,"generated_data/macroabu_firstoccur.RDS")
saveRDS(microabu_firstoccur,"generated_data/microabu_firstoccur.RDS")
```

Save prolonged low eGFR first occurrence table
```{r}
saveRDS(prolonged_low_egfr_firstoccur,"generated_data/prolonged_low_egfr_firstoccur.RDS")
```









