# Phenotype outcomes using combination of tables {#phenotype-outcome-combine}

We achieve the following:

- Generate diabetes first occurrence and multiple event tables by combining:
  - diabetes event tables obtained from the UKB assessment center data (`phenotype_outcome_events_ukb.Rmd`)
  - diabetes event tables obtained from the primary care data (`phenotype_outcome_events_pcp.Rmd`)
  
- Generate diabetic eye disease first occurrence event table by combining:
  - diabetes event tables obtained from the UKB assessment center data  (`phenotype_outcome_events_ukb.Rmd`)
  - diabetes event tables obtained from primary care data (`phenotype_outcome_events_pcp.Rmd`)
  
- Generate advanced chronic kidney disease or diabetic kidney disease first occurrence event table by combining:
  - diabetic kidney disease event table obtained from the UKB assesment center data (`phenotype_outcome_events_ukb.Rmd`)
  - kidney disease case multiple event table obtained from the primary care data (`phenotype_outcome_events_pcp.Rmd`)
  - low creatinine level event table obatained from the biomarker data (`compute_egfr.Rmd` in `phenotype_biomarkers/compute_egfr`)
  - macroalbuminuria first occurrence event table obtained from the biomarker data (`compute_acr_albuminuria.Rmd` in `phenotype_biomarkers/compute_acr_albuminuria`)
  
Input and output files:

- Input:

   - `dm_pc.RDS`
   - `dm_ukb.RDS`
   - `dr_pc.RDS`
   - `dr_firstoccur_ukb.RDS`
   - `kidney_disease_case_pc.RDS`
   - `dkd_ukb.RDS`
   - `low_comb_creat.RDS`
   - `micro_macro_albuminuria_event_tab.RDS`
  
- Output:

  - `dm.RDS`
  - `dm_firstoccur.RDS`
  - `dr_firstoccur.RDS`
  - `dkd_firstoccur.RDS`
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = T)
```

```{r,message=F}
library(tidyverse)
library(data.table)
```

The following outcomes require event tables originating from UKB assessment center data and primary care data 

- Diabetes (DM)
- Diabetic eye disease (DR)
- Diabetic Kidney Disease (CKD)

In particular, phenotyping CKD requires additional event table obtained from using biomarker data.

## DM

Load the following datasets:

- diabetes multiple event table from PCP data
- diabetes multiple event table from UKB assessment center data

```{r}
dm_pc <- readRDS("generated_data/dm_pc.RDS")
dm_ukb <- readRDS("generated_data/dm_ukb.RDS")
```

Create DM multiple and first occurrence event tables
```{r}
dm <- dm_pc %>% full_join(dm_ukb) %>% distinct() %>% arrange(f.eid,event_dt)
dm_firstoccur <- dm %>% group_by(f.eid) %>% arrange(event_dt) %>% slice(1) %>% ungroup()
```

```{r}
saveRDS(dm,"generated_data/dm.RDS")
saveRDS(dm_firstoccur,"generated_data/dm_firstoccur.RDS")
```

## DR

Load the following datasets:

- diabetic eye disease multiple event table from PCP data
- diabetic eye disease first occurrence event table from UKB assessment center data

```{r}
dr_pc <- readRDS("generated_data/dr_pc.RDS")
dr_firstoccur_ukb <- readRDS("generated_data/dr_firstoccur_ukb.RDS")
```

Create diabetic eye disease first occurrence event table (**we are not generating multiple event table for DR**)
```{r}
dr_firstoccur <- 
  dr_pc %>% full_join(dr_firstoccur_ukb) %>% 
  group_by(f.eid) %>% arrange(event_dt) %>% slice(1) %>% ungroup() 
```

```{r}
saveRDS(dr_firstoccur,"generated_data/dr_firstoccur.RDS")
```

## DKD

Load the following datasets:

- kidney disease case multiple event table from PCP data
- diabetic kidney disease multiple event table from UKB assessment center data
- "prolonged low eGFR" first occurrence event table (see the `phenotype_biomarkers` folder for the definition of the "prolonged low eGFR")
- macroabuminuria first occurrence event table

```{r}
kidney_disease_case_pc <- readRDS("generated_data/kidney_disease_case_pc.RDS")
dkd_ukb <- readRDS("generated_data/dkd_ukb.RDS")
prolonged_low_egfr_firstoccur <- readRDS("generated_data/prolonged_low_egfr_firstoccur.RDS")
macroabu_firstoccur <- readRDS("generated_data/macroabu_firstoccur.RDS")
```

Next, we create the following event tables using `low_egfr_firstoccur` and `macroabu_firstoccur` to capture additional DKD cases.
```{r}
prolonged_low_egfr_firstoccur <- prolonged_low_egfr_firstoccur %>% filter(!is.na(event_dt))
macroabu_firstoccur <- macroabu_firstoccur %>% filter(!is.na(event_dt)) %>% select(f.eid,event_dt)
```

Finally, we merge the following event tables to generate first occurrence diabetic kidney disease event table:

- kidney disease case multiple event table from PCP data
- diabetic kidney disease multiple event table from UKB assessment center data
- "prolonged low eGFR" first occurrence event table
- macroabuminuria first occurrence event table

```{r}
dkd_firstoccur <- 
  kidney_disease_case_pc %>% 
  full_join(dkd_ukb) %>%
  full_join(prolonged_low_egfr_firstoccur) %>%
  full_join(macroabu_firstoccur) %>%
  filter(!is.na(event_dt)) %>% 
    group_by(f.eid) %>% arrange(event_dt) %>% slice(1) %>% ungroup()
```

```{r}
saveRDS(dkd_firstoccur,"generated_data/dkd_firstoccur.RDS")
```
