# Generate time-to-event data for macrovascular diabetes complications {#tte-macro}

In this chapter, we phenotype the following macrovascular diabetes compliations:

- Myocardial Infarction (MI)
- Unstable Angina (UA)
- Ischemic Stroke (IS)
- Hemorrhagic Stroke (HS)
- Stroke (ST)
- PCI
- Composite CVD

The following table shows macrovascular complications and associated control exclusion criteria:

|Macrovascular complication|Control exclusion outcome|
|------------|---|
|Myocardial Infarction|cardiovascular control exclusion events|
|Unstable Angina||
|Ischemic Stroke|cerebrovascular control exclusion events|
|Hemorrhagic Stroke|cerebrovascular control exclusion events|
|Stroke|cerebrovascular control exclusion events|
|PCI|non-coronary revascularization procedure control exclusion events|
|Composite CVD| control exclusion events associated with MI, Ischemic stroke, unstable angina and PCI|

The event tables for all of the above control exclusion events were created in chapter \@ref(phenotype-outcomes-ukb). Applying these exclusion criteria to initial time-to-event data gives us the final time-to-event data for macrovascular diabetic complications. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

Load packages and functions.
```{r, message=F}
library(tidyverse)
library(data.table)
source("functions.R")
```

Load first occurrence DM data.
```{r}
dm_firstoccur <- readRDS("generated_data/dm_firstoccur.RDS")
```

Load demographic data.
```{r}
demog <- readRDS("generated_data/demog_selected.RDS")
```

__MI__

Load MI first occurrence data phenotyped from UKB assessment data.
```{r}
mi_firstoccur_ukb <- readRDS("generated_data/mi_firstoccur_ukb.RDS") %>% select(f.eid,event_dt)
```

Load cardiovascular event table that will be used to exclude controls.
```{r}
cardio <- readRDS("generated_data/cardio_control_exclusion_events_ukb.RDS")
```

Define control exclusion subject IDs.
```{r}
ctrl_exclusion_ids_cardio <- cardio$f.eid %>% unique
```

Phenotype time-to-event.
```{r}
mi_tte <- phenotype_time_to_event(dm_firstoccur, mi_firstoccur_ukb,
                        demog, control_exclusion_ids = ctrl_exclusion_ids_cardio)
```

```{r}
saveRDS(mi_tte,"generated_data/mi_tte.RDS")
```

__Unstable angina__

Load unstable angina first occurrence data phenotyped from UKB assessment data.
```{r}
unstable_angina_firstoccur_ukb <- 
  readRDS("generated_data/unstable_angina_firstoccur_ukb.RDS") %>% select(f.eid,event_dt)
```

Phenotype time-to-event.
```{r}
unstable_angina_tte <- phenotype_time_to_event(dm_firstoccur,unstable_angina_firstoccur_ukb,demog)
```

```{r}
saveRDS(unstable_angina_tte,"generated_data/unstable_angina_tte.RDS")
```

__Ischemic stroke__

Load ischemic stroke first occurrence data phenotyped from UKB assessment data.
```{r}
ischemic_stroke_firstoccur_ukb <- readRDS("generated_data/stroke_infarct_firstoccur_ukb.RDS")
```

Load cerebrovascular event table that will be used to exclude controls.
```{r}
cerebro <- readRDS("generated_data/cerebro_control_exclusion_events_ukb.RDS")
```

Define control exclusion subject IDs.
```{r}
ctrl_exclusion_ids_cerebro <- cerebro$f.eid %>% unique
```

Phenotype time-to-event.
```{r}
ischemic_stroke_tte <- phenotype_time_to_event(dm_firstoccur,ischemic_stroke_firstoccur_ukb,demog,
                                               control_exclusion_ids = ctrl_exclusion_ids_cerebro)
```

```{r}
saveRDS(ischemic_stroke_tte,"generated_data/ischemic_stroke_tte.RDS")
```

__Hemorrhagic Stroke__

Load hemorrhagic stroke first occurrence data phenotyped from UKB assessment data.
```{r}
hem_stroke_firstoccur_ukb <- readRDS("generated_data/stroke_hem_firstoccur_ukb.RDS")
```

Load cerebrovascular event table that will be used to exclude controls.
```{r}
cerebro <- readRDS("generated_data/cerebro_control_exclusion_events_ukb.RDS")
```

Define control exclusion subject IDs.
```{r}
ctrl_exclusion_ids_cerebro <- cerebro$f.eid %>% unique
```

Phenotype time-to-event.
```{r}
hem_stroke_tte <- phenotype_time_to_event(dm_firstoccur,hem_stroke_firstoccur_ukb,demog,
                                          control_exclusion_ids = ctrl_exclusion_ids_cerebro)
```

```{r}
saveRDS(hem_stroke_tte,"generated_data/hem_stroke_tte.RDS")
```

__Stroke__

Load stroke first occurrence data phenotyped from UKB assessment data.
```{r}
stroke_firstoccur_ukb <- readRDS("generated_data/stroke_firstoccur_ukb.RDS")
```

Load cerebrovascular event table that will be used to exclude controls.
```{r}
cerebro <- readRDS("generated_data/cerebro_control_exclusion_events_ukb.RDS")
```

Define control exclusion subject IDs.
```{r}
ctrl_exclusion_ids_cerebro <- cerebro$f.eid %>% unique
```

Phenotype time-to-event.
```{r}
stroke_tte <- phenotype_time_to_event(dm_firstoccur,stroke_firstoccur_ukb,demog,
                                      control_exclusion_ids = ctrl_exclusion_ids_cerebro)
```

```{r}
saveRDS(stroke_tte,"generated_data/stroke_tte.RDS")
```

__PCI__

Load PCI first occurrence data phenotyped from UKB assessment data.
```{r}
pci_firstoccur_ukb <- readRDS("generated_data/pci_firstoccur_ukb.RDS")
```

Load non-coronary revascularization event table that will be used to exclude controls.
```{r}
other_revas <- readRDS("generated_data/other_revas_control_exclusion_events_ukb.RDS")
```

Define control exclusion subject IDs.
```{r}
ctrl_exclusion_ids_other_revas <- other_revas$f.eid %>% unique
```

Phenotype time-to-event.
```{r}
pci_tte <- phenotype_time_to_event(dm_firstoccur,pci_firstoccur_ukb,demog,
                                   control_exclusion_ids = ctrl_exclusion_ids_other_revas)
```

```{r}
saveRDS(pci_tte,"generated_data/pci_tte.RDS")
```

__Composite CVD (MI, Unstable Angina, Ischemic Stroke, PCI or CVD death)__

Load composite CVD first occurrence data phenotyped from UKB assessment data.
```{r}
cvd_firstoccur_ukb <- readRDS("generated_data/cvd_firstoccur_ukb.RDS")
```

Create control exclusion table for CVD. They are combination of cardio, cerebro and other revasculrization events that were used to exclude controls when phenotyping MI, unstable angina, ischemic sstroke and PCI time-to-event data.
```{r}
composite_cvd_control_exclusion_events <- list(cardio,cerebro,other_revas) %>% bind_rows()
```

Define control exclusion subject IDs.
```{r}
ctrl_exclusion_ids_cvd <- composite_cvd_control_exclusion_events$f.eid %>% unique
```

Phenotype time-to-event.
```{r}
cvd_tte <- phenotype_time_to_event(dm_firstoccur,cvd_firstoccur_ukb,demog,
                                   control_exclusion_ids = ctrl_exclusion_ids_cvd)
```

```{r}
saveRDS(cvd_tte,"generated_data/cvd_tte.RDS")
```









