# Generate time-to-event data for diabetic eye disease {#tte-dr}

Diabetes first occurrence event table (`dm_firstoccur.RDS`) and diabetic eye disease first occurrence event table (`dr_firstoccur.RDS`) were generated in `phenotype_outcome_events_combine`. Demographic data (`demog_selected.RDS`) was generated in `prep_ukb_assessment`.

For diabetic eye disease, we apply the following exclusion criteria for controls and cases:

- controls are excluded if:
  - represented in UKB diabetic eye disease control exclusion event table 
  - represented in non-diabetic eye disease event table

- controls must satisfy the following to not be excluded:
  - represented in PCP data
  
Subject IDs that were represented in the primary care data (`gp_subject_ids.RDS`) was created in `prep_pcp_data`. The following table shows control exclusion event table files and where they were generated:

|Control exclusion outcome|file|folder|
|------------|---|---|
|UKB eye disease control exclusion event table|`dr_control_exclusion_events_ukb.RDS`|`phenotype_events_ukb`|
|Non-diabetic eye disease event table|`nondm_eye_disease_pc.RDS`|`phenotype_events_pcp`|

Applying these exclusion criteria to initial time-to-event data gives us the final time-to-event data for diabetic eye disease. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, eval = F)
```

```{r}
library(tidyverse)
library(data.table)
source("functions.R")
```

Load the following data sets:

- diabetes first occurrence data
- selected demographic data
- first occurrence DR event table

```{r}
dm_firstoccur <- readRDS("generated_data/dm_firstoccur.RDS")
demog <- readRDS("generated_data/demog_selected.RDS")
dr_firstoccur <- readRDS("generated_data/dr_firstoccur.RDS")
```

Load the following control exclusion event tables:

- DR control exclusion event table from UKB assessment center data
- non-diabeetic eye disease control exclusion event table from PCP data

```{r}
dr_control_exclusion_events_ukb <- readRDS("generated_data/dr_control_exclusion_events_ukb.RDS")
nondm_eye_disease <- readRDS("generated_data/nondm_eye_disease_pc.RDS")
```

Merge control exclusion event tables
```{r}
dr_control_exclusion_events <- full_join(dr_control_exclusion_events_ukb,nondm_eye_disease,by=c("f.eid","event_dt"))
```

Define control exclusion subject IDs
```{r}
ctrl_exclusion_ids <- dr_control_exclusion_events$f.eid %>% unique
```

The controls should be represented in the primary care dataset
```{r}
gp_subejct_ids <- readRDS("generated_data/gp_subject_ids.RDS")
```

Define control inclusion subject IDs
```{r}
ctrl_inclusion_ids <- gp_subejct_ids %>% unique
```

Phenotpe DR time-to-event table
```{r}
dr_tte <- phenotype_time_to_event(dm_firstoccur,dr_firstoccur,demog,
                                  control_exclusion_ids = ctrl_exclusion_ids,
                                  control_inclusion_ids = ctrl_inclusion_ids)
```

```{r}
saveRDS(dr_tte,"generated_data/dr_tte.RDS")
```





