# Generate time-to-event data for diabetic eye disease {#tte-dr}

In this chapter, we phenotype right-censored time-to-event data for diabetic eye disease. An initial time-to-event data are created, and consequent cases and controls are identified following the identical procedure described in chapter \@ref(tte-macro). The initial time-to-event data are refined by applying the following exclusion criteria: 

  - case
    - has some event represented in diabetic eye disease control exclusion event table (generated in \@ref(phenotype-outcomes-ukb))
    - has some event represented in non-diabetic eye disease event table (generated in \@ref(phenotype-outcomes-pcp))
    - has a complication event after the censoring date
    - has a prior complication event identified before a first evidence of diabetes
    - has an initiation date that is not between the first evidence of diabetes and the date of complication event
  
  - control
    - has a first evidence of diabetes after the censoring date
    - has less than 5 years of follow-up time since the first evidence of diabetes before censoring date
    - has a first evidence of diabetes 6 months after the study initiation date
    - is not represented in the primary care data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, eval = F)
```

Load packages and functions.
```{r}
library(tidyverse)
library(data.table)
source("functions.R")
```

Load the following datasets:

- diabetes first occurrence data
- selected demographic data
- first occurrence DR event table

```{r}
dm_firstoccur <- readRDS("generated_data/dm_firstoccur.RDS")
demog <- readRDS("generated_data/demog_selected.RDS")
dr_firstoccur <- readRDS("generated_data/dr_firstoccur.RDS")
```

Load the following control exclusion event tables:

- DR control exclusion event table from UKB assessment center data
- non-diabeetic eye disease control exclusion event table from the primary care data

```{r}
dr_control_exclusion_events_ukb <- readRDS("generated_data/dr_control_exclusion_events_ukb.RDS")
nondm_eye_disease <- readRDS("generated_data/nondm_eye_disease_pc.RDS")
```

Merge control exclusion event tables.
```{r}
dr_control_exclusion_events <- full_join(dr_control_exclusion_events_ukb,nondm_eye_disease,by=c("f.eid","event_dt"))
```

Define control exclusion subject ID's.
```{r}
ctrl_exclusion_ids <- dr_control_exclusion_events$f.eid %>% unique
```

The controls should be represented in the primary care data.
```{r}
gp_subejct_ids <- readRDS("generated_data/gp_subject_ids.RDS")
```

Define control inclusion subject ID's.
```{r}
ctrl_inclusion_ids <- gp_subejct_ids %>% unique
```

Phenotype time-to-event table for diabetic eye disease.
```{r}
dr_tte <- phenotype_time_to_event(dm_firstoccur,dr_firstoccur,demog,
                                  control_exclusion_ids = ctrl_exclusion_ids,
                                  control_inclusion_ids = ctrl_inclusion_ids)
```

```{r}
saveRDS(dr_tte,"generated_data/dr_tte.RDS")
```





