# Generate time-to-event data for diabetes related kidney disease {#tte-dkd}

Diabetes first occurrence event table (`dm_firstoccur.RDS`) and diabetic kidney disease first occurrence event table (`dkd_firstoccur.RDS`) were generated in `phenotype_outcome_events_combine`. Demographic data (`demog_selected.RDS`) was generated in `prep_ukb_assessment`.

For CKD, we apply the following exclusion criteria for controls and cases:

- controls are excluded if:
  - represented in UKB CKD control exclusion event table 
  - represented in PCP CKD control exclusion event table
  - found to have microalbuminuria at least once
  - found to have eGFR level < 60

- controls must satisfy all of the following to not be excluded:
  - represented in PCP data
  - have eGFR measurements available

- cases were excluded if:
  - the follow-up time was less than 5 years
  
Subject IDs that were represented in the primary care data (`gp_subject_ids.RDS`) was created in `prep_pcp_data`. The control exclusion based on microalbuminuria status and the eGFR measurements were accomplished based on corresponding biomarker datasets `micro_macro_albuminuria_event_tab.RDS` and `comb_creat_tab.RDS`, respectively, both generated in `phenotype_biomarkers` folder. The following table shows control exclusion event table files and where they were generated:

|Control exclusion outcome|file|folder|
|------------|---|---|
|UKB CKD control exclusion event table|`dkd_control_exclusion_events_ukb.RDS`|`phenotype_events_ukb`|
|PCP CKD control exclusion event table|`kidney_disease_control_exclusion_pc.RDS`|`phenotype_events_pcp`|
|microalbuminuria status|`micro_macro_albuminuria_event_tab.RDS`|`phenotype_biomarkers/compute_acr_albuminuria`|
|eGFR measurements|`comb_creat_tab.RDS`|`phenotype_biomarkers/compute_egfr`|
  
Applying these exclusion criteria to initial time-to-event data gives us the final time-to-event data for diabetic kidney disease. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, eval = F)
```

```{r}
library(tidyverse)
library(data.table)
source("functions.R")
```

Load the following datasets:

- diabetes first occurrence data
- selected demographic data
- first occurrence DKD event table
- eGFR trajectory data
- microalbuminuria first occurrence event table

```{r}
dm_firstoccur <- readRDS("generated_data/dm_firstoccur.RDS")
demog <- readRDS("generated_data/demog_selected.RDS")
dkd_firstoccur <- readRDS("generated_data/dkd_firstoccur.RDS")
egfr_traj <- readRDS("generated_data/trajectory_egfr.RDS")
microabu_firstoccur <- readRDS("generated_data/microabu_firstoccur.RDS")
```

Load the following control exclusion event tables:

- CKD control exclusion event table from UKB assessment center data
- CKD control exclusion event table from PCP data

```{r}
dkd_control_exclusion_events_ukb <- readRDS("generated_data/dkd_control_exclusion_events_ukb.RDS")
kidney_disease_control_exclusion_events_pc <- readRDS("generated_data/kidney_disease_control_exclusion_pc.RDS")
```

Create additional control exclusion event tables:

- a table containing subjects who have been recorded to have ever had creatinine level less than 60 
- a table containing subjects who have been recorded to have ever had microalbuminuria

```{r}
egfr_lt60_ever_event_tab <- egfr_traj %>% filter(measurement < 60, !is.na(measurement), !is.na(event_dt))
microabu_ever_event_tab <- microabu_firstoccur %>% filter(!is.na(event))
```

Merge control exclusion event tables
```{r}
ckd_control_exclusion_events <- bind_rows(list(dkd_control_exclusion_events_ukb,
            kidney_disease_control_exclusion_events_pc,
            egfr_lt60_ever_event_tab,microabu_ever_event_tab))
```

Define control exclusion subject IDs
```{r}
ctrl_exclusion_ids <- ckd_control_exclusion_events$f.eid %>% unique
```

The controls should have linked primary care dataset and at least one eGFR measurement
```{r}
pcp_subejct_ids <- readRDS("generated_data/gp_subject_ids.RDS")
eGFR_avail_subject_ids <- (egfr_traj %>% filter(!is.na(measurement) & !is.na(event_dt))) %>% .$f.eid
```

Define control inclusion subject IDs
```{r}
ctrl_inclusion_ids <- intersect(pcp_subejct_ids,eGFR_avail_subject_ids) %>% unique
```

```{r}
dkd_tte <- phenotype_time_to_event(dm_firstoccur,dkd_firstoccur,demog,
                                  control_exclusion_ids = ctrl_exclusion_ids,
                                  control_inclusion_ids = ctrl_inclusion_ids)
```

We filter out cases that do not have 5+ follow-up time
```{r}
dkd_tte <- 
  dkd_tte %>%
  filter(event == 0 | (event == 1 & (lubridate::decimal_date(event_dt_comp) - lubridate::decimal_date(event_dt_dm) >= 5)))
```

```{r}
saveRDS(dkd_tte,"generated_data/dkd_tte.RDS")
```


